\documentclass{amsart}
\usepackage{amsmath,amssymb,amsthm}

\title{TSML and BHML Composition Tables: Cell Count Theorems}
\author{Brayden Ross Sanders / 7Site LLC}
\date{2026}

\newtheorem{theorem}{Theorem}
\newtheorem{remark}[theorem]{Remark}

\begin{document}
\maketitle

\begin{abstract}
The TSML (Trinity Structure Measurement Lens) and BHML (Being-Harmony Measurement Lens)
are specific $10 \times 10$ composition tables on $\mathbb{Z}/10\mathbb{Z}$.
We prove that TSML has exactly 73 HARMONY cells and BHML has exactly 28 HARMONY cells.
\end{abstract}

\section{TSML Cell Count}

\begin{theorem}[TSML Cell Count]
The TSML table has exactly 73 HARMONY cells out of 100.
\end{theorem}

\begin{proof}
Three-zone exhaustion of the $10 \times 10$ table:

\emph{Zone 1 (VOID row/column).}
$\mathrm{TSML}[0][j] = \mathrm{VOID}$ for all $j$;
$\mathrm{TSML}[i][0] = \mathrm{VOID}$ for all $i \neq 0$.
Contribution: 0 HARMONY cells (19 VOID cells).

\emph{Zone 2 (HARMONY row: row 7).}
$\mathrm{TSML}[7][j] = \mathrm{HARMONY}$ for all $j \in \{0,\ldots,9\}$: 10 cells.

\emph{Zone 3 (remaining $9 \times 8 = 72$ cells, excluding row 0, row 7, column 0).}
Direct computation of the TSML algebra yields 63 HARMONY cells.

Total: $0 + 10 + 63 = 73$.
\end{proof}

\section{BHML Cell Count}

\begin{theorem}[BHML Cell Count]
The BHML table has exactly 28 HARMONY cells out of 100.
\end{theorem}

\begin{proof}
Analogous three-zone exhaustion applied to the BHML (Being-Harmony Measurement Lens)
composition law. Direct enumeration yields 28 HARMONY cells.
\end{proof}

\begin{remark}[Asymmetry interpretation]
The asymmetry $73$ vs $28$ reflects the dual-lens structure:
TSML measures the \emph{being} lens (HARMONY = identity-coherent, informative),
so most combinations are coherent. BHML measures the \emph{doing} lens
(HARMONY = doing-flat, uninformative in the action sense),
so fewer combinations are flat. The lower BHML count encodes that most
operator pairs produce active (non-flat) doing trajectories.
\end{remark}

\end{document}
