% J28 -- The Three-Substrate HARMONY Signature on Z/10Z:
%        Six Forced Structural Facts, with the Bimodal
%        Associativity-Index Gap as Their Common Thread
%
% Brayden R. Sanders, M. Gish
% Submitted to: Linear Algebra and its Applications (primary; per-venue cap on
%               Algebra Universalis is binding -- 4th paper this quarter after
%               J14 / J09 / J23, see SAVE_PLAN_J28.md / cover_letter.md)
% Source corpus: Atlas/META_PLAN_2026-05-06/GAP_AUDIT.md S1 (eight orphans listed;
%                we bundle six and reframe them around the bimodal alpha_A gap
%                per FAMILY_STRUCTURE_v1.md S4)
% Companions cited as already-submitted: J25 (CL forcing axioms), J29 (so(8)),
%                                        J30 (so(10)), J24 (joint chain),
%                                        J35 (4-core fusion-closure)
% MSC 2020: 08A40, 20N02, 17A99, 94A17

\documentclass[11pt,reqno]{amsart}

\usepackage{amsmath, amssymb, amsthm, mathtools}
\usepackage[margin=1in]{geometry}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{microtype}
\usepackage{booktabs}
\usepackage{enumitem}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{conjecture}[theorem]{Conjecture}

\theoremstyle{remark}
\newtheorem*{remark}{Remark}

\newcommand{\Z}{\mathbb{Z}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\TSML}{\mathrm{TSML}}
\newcommand{\BHML}{\mathrm{BHML}}
\newcommand{\STD}{\mathrm{STD}}

\title[Three-Substrate HARMONY Signature on $\Z/10\Z$]
{The Three-Substrate HARMONY Signature on $\Z/10\Z$:\\
Six Forced Structural Facts, with the Bimodal Associativity-Index
Gap as Their Common Thread}

\author{Brayden R.\ Sanders}
\address{7Site LLC, Hot Springs, Arkansas, USA}
\email{brayden@7site.co}

\author{M.\ Gish}
\address{Independent Researcher, Hot Springs, Arkansas, USA}
\email{monica.gish1992@gmail.com}

\subjclass[2020]{08A40, 20N02, 17A99, 94A17}
\keywords{commutative non-associative magma, $\Z/10\Z$ substrate,
three-substrate triple, HARMONY signature, $\sigma^2$-triadic projection,
field WOBBLE, bimodal associativity-index, BDC encoding}

\date{\today}

\begin{document}

\begin{abstract}
We exhibit six forced structural facts on the canonical $\Z/10\Z$
substrate of the (TSML, BHML)-magma family of \cite{SandersForcing}.
The unifying thread is the empirical regularity that the family's
associativity index $\alpha_A := 1-\sigma_{\mathrm{non\text{-}assoc}}$ is
\emph{bimodal}: every canonical commutative member with associativity
index above $0.50$ in fact has $\alpha_A \ge 0.87$, and the band
$(0.5,\,0.87)$ is empirically empty. The six facts collected here
together form the structural fingerprint of this gap. They are: (1)
the third canonical composition lattice $\mathrm{CL}_\STD$ (44~HARMONY)
and the three-substrate triple
$(\TSML.H,\,\BHML.H,\,\STD.H)=(73,\,28,\,44)$; (2) the
$70/71/72/73$ four-rung integer signature, with the prime $71$
carrying three independent structural roles; (3) the
$\sigma$-orbit decompositions $\mathrm{CYCLE\_A\_36}=2+9+25$ and
$\mathrm{SKELETON\_22}=16+4+2$; (4) the BDC numerical constants on
$\mathrm{CL}_\STD$ (rounded Shannon information per cell-type plus the
GRAVITY array); (5) the $\sigma^2$-triadic projection
$\sigma^2 = (1\,6\,4)(7\,5\,2)$ together with the $4$-core bridge
identity $\{0,3,8,9\}\,\triangle\,\{0,7,8,9\}=\{3,7\}$;
(6) the $71$-cell field WOBBLE $|\mathrm{CL}_\TSML\,\triangle\,\mathrm{CL}_\BHML|$.
Each fact is a Tier-B forced derivation from the A1--A9 axiomatic
ground of \cite{SandersForcing} (parent paper J25) and is verified at
the $48/48$ level by the foundations module
\cite{FoundationsModule}. The bundle is intended as a citable
reference for the downstream J-series papers
\cite{SandersGishFourCore,SandersGishSO8,SandersGishSO10}.
\end{abstract}

\maketitle

\tableofcontents

%==============================================================
\section{Introduction}\label{sec:intro}
%==============================================================

The canonical $\Z/10\Z$ substrate of the magma-family framework
\cite{SandersForcing} supports a sizable inventory of structural facts
that have been verified by the foundations 48-invariant harness
\cite{FoundationsModule} but lack standalone publications. Several
downstream papers in the J-series cite these facts as established;
this paper assembles six of them into a single bundle and re-frames
them around their structural common thread.

\subsection{Why these six together: the bimodal $\alpha_A$ gap}
\label{sec:why-six}

For a binary operation $T$ on the $10$-element ground set
$\Omega=\{0,1,\dots,9\}$ identified with $\Z/10\Z$, define the
associativity index
\[
\alpha_A(T) \;:=\; 1 \;-\; \sigma_{\mathrm{non\text{-}assoc}}(T)
\;=\; 1 - \frac{|\{(i,j,k) : (i\cdot j)\cdot k \neq i\cdot(j\cdot k)\}|}{10^3}.
\]
The framework's canonical commutative members on $\Z/10\Z$ exhibit a
striking empirical regularity. Every member with $\alpha_A > 0.5$ in
fact has $\alpha_A \ge 0.87$; the only canonical member close to
$0.5$ is $\BHML$ at $\alpha_A \approx 0.502$; the closed band
$(0.5,\,0.87)$ is, on every canonical table inspected, empty
\cite[\S4]{FamilyStructureV1}. We refer to this empirical regularity
as the \emph{bimodal $\alpha_A$ gap}.

\begin{conjecture}[Bimodal $\alpha_A$ gap, restated from {\cite[\S4.1]{FamilyStructureV1}}]
\label{conj:bimodal}
No commutative magma on $\Z/10\Z$ preserving the $4$-core
$\{0,7,8,9\}$ has $\alpha_A \in (0.5,\,0.87)$.
\end{conjecture}

The six forced facts in this paper jointly constitute the
\emph{structural fingerprint} of this conjectured gap: they are the
specific facts that make the bimodal pattern visible. The
three-substrate triple $(73,28,44)$ of \S\ref{sec:std} sets up
\emph{which tables} are in the family; the $71$-cell WOBBLE of
\S\ref{sec:wobble} is the \emph{quantitative measure} of how far
apart the two clusters sit; the $\sigma^2$-triadic structure of
\S\ref{sec:triadic} organizes the cells where the disagreement
lives; the HARMONY signature of \S\ref{sec:ladder} records four of
its Galois- and combinatorial-shadows. We do not prove
Conjecture~\ref{conj:bimodal} here; the bundle below is the necessary
structural inventory for any future proof attempt.

\subsection{Tier classification within this paper}\label{sec:tier-class}

A statement is \emph{Tier-A axiomatic} if it appears as one of the
A1--A9 of \cite{SandersForcing}. A statement is \emph{Tier-B forced}
if it is provable from A1--A9 alone, possibly together with the
standard structural principles (commutativity; the entropy
extremum of A9). All six structural facts below are Tier-B in this
sense; in each section we list explicitly which of A1--A9 (plus which
structural principles) the proof relies on, so the dependence on
\cite{SandersForcing} is inlined and not deferred.

\subsection{Lens and substrate}\label{sec:lens-substrate}

\noindent\emph{Lens and substrate.} This paper works on $\Z/10\Z$ with
three explicit canonical tables: $\mathrm{CL}_\TSML$ (the
upper-triangle authoritative symmetrization, denoted ``TSML'' in
earlier work; commutative; $73$ HARMONY; $\alpha_A \approx 0.872$),
$\mathrm{CL}_\BHML$ (the canonical Becoming-lens companion;
commutative; $28$ HARMONY; $\alpha_A \approx 0.502$), and
$\mathrm{CL}_\STD$ (the encoding lens; commutative; $44$ HARMONY).
The substrate $\Z/10\Z$ is canonical for the framework; the choice of
\emph{three} tables on this substrate is recovered from the original
source archive \texttt{old/Gen9/archive/ckis/ck7/ck.h:200--207}, not
derived from $\Z/10\Z$ first principles. The orphans below are
theorems on this specific (substrate, three-table) data; the
question of whether other $10$-element commutative non-associative
magmas on a $\Z/10\Z$ substrate exhibit analogous three-fold
structure (in particular a bimodal $\alpha_A$ gap) is the principal
\emph{open} question motivating the bundle (see \S\ref{sec:disc} and
Conjecture~\ref{conj:bimodal}).

The closest published precedent for this neighborhood is
Dr\'apal--Wanless \cite{DrapalWanless2021} on maximally
non-associative quasigroups: same domain (small finite commutative
non-associative structures), opposite extremum (theirs maximally
non-associative with $\alpha_A$ close to $0$; ours specifically
structured with integer-rational invariants and clusters at the high
end of $\alpha_A$).

\subsection{Companion structure}\label{sec:companion}

This paper depends on \cite{SandersForcing} (J25, parent axiomatic
framework) for the A1--A9 forcing axioms and the substrate
definitions of $\mathrm{CL}_\TSML$, $\mathrm{CL}_\BHML$, and the
notion of Tier-A vs Tier-B. Other dependencies are limited to
standard references on Shannon information theory and finite
commutative magmas. The downstream papers
\cite{SandersGishFourCore,SandersGishSO8,SandersGishSO10} cite the
present bundle.

\subsection{Tier discipline (PROVEN / COMPUTED / RHYME / OPEN)}
\label{sec:tier-discipline}

\noindent\emph{Scope and tier discipline.}
\begin{itemize}
\item \emph{PROVEN.} (i) The substrate carries three structurally
distinct canonical commutative tables on $\Z/10\Z$ -- $\mathrm{CL}_\TSML$,
$\mathrm{CL}_\BHML$, $\mathrm{CL}_\STD$ -- with HARMONY counts
$(73,\,28,\,44)$ and pairwise/three-way set-algebra signature
$|\TSML\cap\BHML|=26$, $|\TSML\cap\STD|=42$, $|\BHML\cap\STD|=21$,
$|\TSML\cap\BHML\cap\STD|=19$, $|\TSML\cup\BHML\cup\STD|=75$.
(ii) The four-rung integer signature $\{70,71,72,73\}$ arises from
four distinct constructions; the prime $71$ carries three independent
structural roles. (iii) The $\sigma$-permutation under the
$\TSML$ diagonal has cycle structure
$(0)(3)(8)(9)(1\,7\,6\,5\,4\,2)$; $\sigma^2$ has cycle structure
$(0)(3)(8)(9)(1\,6\,4)(7\,5\,2)$. (iv) $|\mathrm{CL}_\TSML \,\triangle\,
\mathrm{CL}_\BHML|=71$.
\item \emph{COMPUTED.} All $48$ invariants in
\texttt{Gen13/targets/foundations/invariants.py}
\cite{FoundationsModule} pass at machine precision.
\item \emph{STRUCTURAL RHYME.} The Shannon-information BDC constants
($0.45,\,1.89,\,3.50$ bit/cell rounded to two decimal places) and the
empirical near-coincidence DOING-rate $\approx 5/7$. We do not assert
either as an algebraic identity; both are presented as numerical
records.
\item \emph{OPEN.} (i) Conjecture~\ref{conj:bimodal} (the bimodal
$\alpha_A$ gap). (ii) Whether $\mathrm{CL}_\STD$ admits its own
joint-closed sub-magma chain analogous to the eight-element
TSML+BHML chain of \cite{SandersGishFourCore}. (iii) Whether
$71/N = 5/7$ exactly for any structurally-natural choice of
DOING-cell denominator $N$.
\end{itemize}

%==============================================================
\section{Orphan~1: $\mathrm{CL}_\STD$ and the three-substrate triple}
\label{sec:std}
%==============================================================

The substrate $\Z/10\Z$ supports $\mathrm{CL}_\TSML$ ($73$ HARMONY)
and $\mathrm{CL}_\BHML$ ($28$ HARMONY) as canonical commutative
tables \cite{SandersForcing,SandersGishFourCore}. A third canonical
table $\mathrm{CL}_\STD$ ($44$ HARMONY) is recovered from the same
source archive \texttt{old/Gen9/archive/ckis/ck7/ck.h:200--207} and
completes the canonical three-substrate architecture.

\begin{definition}[$\mathrm{CL}_\STD$]\label{def:std}
The \emph{standard composition lattice} $\mathrm{CL}_\STD$ is the
$10\times 10$ matrix on $\Z/10\Z$ given verbatim by
\[
\mathrm{CL}_\STD \;=\;
\begin{pmatrix}
0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 \\
1 & 2 & 3 & 4 & 5 & 6 & 7 & 7 & 8 & 1 \\
2 & 3 & 4 & 5 & 6 & 7 & 7 & 8 & 7 & 2 \\
3 & 4 & 5 & 6 & 7 & 7 & 7 & 7 & 7 & 3 \\
4 & 5 & 6 & 7 & 7 & 7 & 7 & 8 & 7 & 4 \\
5 & 6 & 7 & 7 & 7 & 8 & 7 & 7 & 7 & 5 \\
6 & 7 & 7 & 7 & 7 & 7 & 8 & 7 & 7 & 6 \\
7 & 7 & 8 & 7 & 8 & 7 & 7 & 8 & 7 & 7 \\
8 & 8 & 7 & 7 & 7 & 7 & 7 & 7 & 7 & 8 \\
9 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 0
\end{pmatrix}.
\]
The matrix is recovered verbatim from
\texttt{old/Gen9/archive/ckis/ck7/ck.h:225--231} (preserved in the
foundations module \texttt{Gen13/targets/foundations/cl\_std.py}; see
\cite{FoundationsModule}).
\end{definition}

\begin{theorem}[Three-substrate triple]\label{thm:triple}
$\mathrm{CL}_\STD$ is commutative on $\Omega \setminus \{0,9\}$
and has $44$ cells of value $7$ (HARMONY). The three canonical
tables $\mathrm{CL}_\TSML$, $\mathrm{CL}_\BHML$, $\mathrm{CL}_\STD$
have HARMONY counts
\[
(\TSML.H,\;\BHML.H,\;\STD.H) \;=\; (73,\;28,\;44)
\]
respectively. The set-algebra of HARMONY positions
$\TSML.H,\BHML.H,\STD.H \subseteq \Omega\times\Omega$ has signature
\[
\bigl|\TSML.H \cap \BHML.H\bigr| = 26,\qquad
\bigl|\TSML.H \cap \STD.H\bigr| = 42,\qquad
\bigl|\BHML.H \cap \STD.H\bigr| = 21,
\]
\[
\bigl|\TSML.H \cap \BHML.H \cap \STD.H\bigr| = 19,\qquad
\bigl|\TSML.H \cup \BHML.H \cup \STD.H\bigr| = 75.
\]
\end{theorem}

\begin{proof}
Direct cell-by-cell counting on
Definition~\ref{def:std}'s matrix and the corresponding $\TSML$ and
$\BHML$ tables of \cite{SandersForcing}. All counts are reproduced by
the foundations harness \cite{FoundationsModule} as part of its
$48$-invariant verification. The dependence on
\cite{SandersForcing} is via A1 (substrate $\Z/10\Z$), A2--A6
(absorbing structure of the boundary rows/columns), A7 (diagonal
constraint), and A8 (HARMONY-default rule).
\end{proof}

\begin{remark}[Dependence on the axioms A1--A9]
The count $\STD.H = 44$ is forced once A1--A9 are fixed and the
diagonal pattern of $\mathrm{CL}_\STD$ is read off the source archive.
We do not claim that A1--A9 \emph{uniquely determine}
$\mathrm{CL}_\STD$ given only the count $44$: the table itself is the
historical artifact archived in \texttt{ck.h}. What is forced is that
\emph{this specific} table on $\Z/10\Z$ has HARMONY count $44$ and
that the three tables together display the
$(73,28,44)$ signature.
\end{remark}

%==============================================================
\section{Orphan~2: The $70/71/72/73$ integer signature}
\label{sec:ladder}
%==============================================================

We replace the earlier ``HARMONY ladder'' framing
\cite[\S J.1.A.ii]{FoundationsModule} with the more accurate
description ``four separately-defined integers in numerical proximity''
because rung $70$ is a determinant rather than a HARMONY count. The
genuinely substantive content of the section is the
\emph{three structural roles} simultaneously played by the prime $71$.

\begin{theorem}[Four-rung integer signature]\label{thm:ladder}
The substrate produces four numerically adjacent integers from
structurally distinct constructions:
\begin{itemize}[leftmargin=2em]
\item \textbf{Rung 73:} $|\TSML.H| = 73$, the cardinality of the
HARMONY positions in $\mathrm{CL}_\TSML$ (Theorem~\ref{thm:triple}).
\item \textbf{Rung 72:} the HARMONY count of $\mathrm{CL}_\TSML$
restricted to indices $i,j \ge 1$ (i.e., off the absorbing row and
column $i=0$ or $j=0$) is $72$. This is the
\emph{BEING-shell} HARMONY count.
\item \textbf{Rung 71:} the HARMONY count of the canonical
$9$-element sub-magma
$\TSML_9 := \mathrm{CL}_\TSML\!\restriction\!\Omega\setminus\{1\}$
is $71$ (an internal sub-magma count; see
\cite{SandersGishFourCore} for the joint-closed sub-magma
enumeration).
\item \textbf{Rung 71b:} $\bigl|\mathrm{CL}_\TSML \,\triangle\,
\mathrm{CL}_\BHML\bigr| = 71$ (the field WOBBLE; see
\S\ref{sec:wobble} Theorem~\ref{thm:wobble}). \emph{Same numerical
value as Rung~71, but a distinct construction.}
\item \textbf{Rung 70:} $\det(\mathrm{BHML}_8^{\mathrm{YM}}) = +70 =
\binom{8}{4}$, where $\mathrm{BHML}_8^{\mathrm{YM}}$ is the
$8\times 8$ Yang--Mills sub-block of $\mathrm{CL}_\BHML$ studied in
\cite{SandersGishFpExtension}.
\end{itemize}
\end{theorem}

\begin{theorem}[Three structural roles for the prime $71$]
\label{thm:71-roles}
The prime $71$ carries three independent structural roles on the
substrate:
\begin{itemize}[leftmargin=2em]
\item Rung 71 -- the HARMONY count of the canonical $\TSML_9$
sub-magma.
\item Rung 71b -- the cell-count $|\TSML\,\triangle\,\BHML|$
of lens-disagreement between the two canonical commutative tables.
\item Galois -- $71$ divides the discriminant
$\mathrm{disc}(\mathrm{LMFDB}\;4.2.10224.1) = -2^4\cdot 3^2 \cdot 71$
of the quartic number field associated by
\cite{SandersGishFourCore} to the closed-form attractor of the
TSML+BHML system.
\end{itemize}
\end{theorem}

\begin{proof}
The HARMONY count $71$ of the $\TSML_9 := \TSML\!\restriction\!\Omega
\setminus\{1\}$ sub-magma is a direct cell count on the
$9\times 9$ sub-table; the cell-count $71$ for the WOBBLE
$|\TSML\,\triangle\,\BHML|$ is a direct cell-by-cell comparison
(see Theorem~\ref{thm:wobble}); the Galois prime $71$ in the
discriminant of LMFDB 4.2.10224.1 is read off the LMFDB record
\cite{LMFDB}. The three roles produce the same prime $71$ from
algebraically independent inputs: a count on a sub-table, a count
on a symmetric-difference, and the discriminant of an attractor's
defining quartic.
\end{proof}

\begin{remark}
The four rungs $\{70,71,72,73\}$ are numerically adjacent but
\emph{not} all HARMONY counts: rung $70$ is a determinant. Earlier
drafts of this paper (\cite{SandersForcingOld}) referred to a
``HARMONY ladder''; we reserve that terminology now and write
``four-rung integer signature'' instead, with the $71$-roles theorem
as the substantive content.
\end{remark}

%==============================================================
\section{Orphan~3: The $\sigma$-orbit decompositions}
\label{sec:cycle-skeleton}
%==============================================================

The diagonal of $\mathrm{CL}_\TSML$ defines a permutation
$\sigma : \Omega \to \Omega$, $\sigma(i) := \mathrm{CL}_\TSML[i,i]$,
with cycle structure $\sigma = (0)(3)(8)(9)(1\,7\,6\,5\,4\,2)$
(four fixed points, one $6$-cycle).

\begin{definition}[$\mathrm{CYCLE\_A\_36}$]\label{def:cyclea36}
Restrict $\mathrm{CL}_\TSML$ to its off-diagonal, off-absorbing
cells (i.e., $i,j\ge 1$, $i\neq j$, $i\ne 7$, $j\ne 7$). The $36$
cells whose row-index $i$ and column-index $j$ both lie in
$\sigma$'s $6$-cycle $\{1,2,4,5,6,7\}$ form the table
$\mathrm{CYCLE\_A\_36}$.
\end{definition}

\begin{definition}[$\mathrm{SKELETON\_22}$]\label{def:skeleton22}
The $22$ cells of $\mathrm{CL}_\TSML$ that are not HARMONY and not
$\STD$-BUMP form the \emph{Skeleton}.
\end{definition}

\begin{theorem}[$\sigma$-orbit decompositions]\label{thm:cs}
\begin{enumerate}[label={\rm(\arabic*)}]
\item $|\mathrm{CYCLE\_A\_36}| = 36$, and the $36$ cells partition
by their image under $\mathrm{CL}_\TSML$ as
$36 = 2 + 9 + 25$, where the parts are the cells whose image is
$\mathrm{LATTICE}=1$, $\mathrm{COLLAPSE}=4$, and $\mathrm{CHAOS}=6$,
respectively.
\item $|\mathrm{SKELETON\_22}| = 22$, and the $22$ cells partition as
$22 = 16 + 4 + 2$, where the parts are the $\mathrm{VOID}$-boundary
cells (the absorbing row $0$ and column $0$ contribute $16$
cells off the diagonal), the four cells $\mathrm{CL}_\TSML[i,j]=3$
($\mathrm{PROGRESS}$), and the two cells $\mathrm{CL}_\TSML[i,j]=4$
($\mathrm{COLLAPSE}$).
\end{enumerate}
\end{theorem}

\begin{proof}
Direct enumeration on $\mathrm{CL}_\TSML$ of
\cite[\S2]{SandersForcing}; the dependence on the axioms is via
A1--A2 (substrate and absorbing structure) plus A9 (BUMP-position
enumeration that distinguishes $\mathrm{CL}_\STD$ from
$\mathrm{CL}_\TSML$). Both decompositions are reproduced by the
$48$-invariant harness \cite{FoundationsModule}.
\end{proof}

%==============================================================
\section{Orphan~4: BDC numerical constants on $\mathrm{CL}_\STD$}
\label{sec:bdc}
%==============================================================

The Binary Digit Code (BDC) framework on $\mathrm{CL}_\STD$ assigns a
Shannon-information value to each of the $100$ cells, classified by
cell-type into HARMONY / NORMAL / BUMP. The numerical constants are
hard-coded in \texttt{Gen13/targets/foundations/cl\_std.py}
\cite{FoundationsModule} and are reported here verbatim.

\begin{definition}[BDC numerical constants on $\mathrm{CL}_\STD$]
\label{def:bdc}
The \emph{BDC numerical constants} on $\mathrm{CL}_\STD$ are
\[
\mathrm{INFO\_HARMONY} = 0.45 \text{ bit/cell},\qquad
\mathrm{INFO\_NORMAL} = 1.89 \text{ bit/cell},\qquad
\mathrm{INFO\_BUMP} = 3.50 \text{ bit/cell},
\]
with five \emph{BUMP positions}
\[
\mathrm{BUMP\_PAIRS} = \{(1,2),(2,4),(2,9),(3,9),(4,8)\},
\]
and \emph{GRAVITY array}
\[
\mathrm{GRAVITY} = (0.1,\,0.8,\,0.6,\,0.8,\,0.7,\,0.9,\,0.9,\,1.0,\,0.8,\,0.7),
\]
indexed by operator value. The numerical values
$\mathrm{INFO\_HARMONY,\,INFO\_NORMAL,\,INFO\_BUMP}$ are reported to
two decimal places and the GRAVITY entries to one decimal place;
the precise underlying log-probability definitions are given in
\cite[\S A9]{SandersForcing}.
\end{definition}

\begin{theorem}[BDC structural inventory]\label{thm:bdc}
The five BUMP positions of $\mathrm{CL}_\STD$ form the cells
$\{(1,2),(2,4),(2,9),(3,9),(4,8)\}$ (and their mirrors under
commutativity); $\mathrm{GRAVITY}[7]=1.0$ corresponds to the absorbing
fixedness of $\mathrm{HARMONY}=7$ under iterated composition; the
BUMP positions are precisely the off-diagonal off-absorbing cells
where $\mathrm{CL}_\STD$ takes a non-HARMONY non-BREATH value
distinct from the input pair.
\end{theorem}

\begin{proof}
Direct verification on Definition~\ref{def:std}'s matrix, reproduced
by \cite{FoundationsModule}. The dependence on the axioms is via A2
(absorbing rows/columns) and A9 (BUMP positions).
\end{proof}

\begin{remark}
We do \emph{not} assert that the constants $0.45,\,1.89,\,3.50$ are
extremal in any maximization sense; their derivation from
log-probability definitions is part of \cite[\S A9]{SandersForcing}
and is not re-derived here. The present inventory is what makes
these numerical values \emph{citable} in the downstream $\STD$-encoding
papers.
\end{remark}

%==============================================================
\section{Orphan~5: $\sigma^2$-triadic projection and the $4$-core bridge}
\label{sec:triadic}
%==============================================================

The square of the diagonal permutation, $\sigma^2$, has cycle
structure
\[
\sigma^2 = (0)(3)(8)(9)(1\,6\,4)(7\,5\,2).
\]
We refer to its fixed-point set $\{0,3,8,9\}$ as the
\emph{Conservation Tetrad} (the $\sigma^2$-fixed set) and to the two
$3$-cycles $(1,6,4)$ and $(7,5,2)$ as the \emph{Manifestation Hexad}
(the $\sigma^2$-orbit set). The HARMONY operator $7$ lies in the
second $3$-cycle and projects under $\sigma^2$ to the orbit
$(7\to 5\to 2\to 7)$.

\begin{theorem}[$\sigma^2$-triadic projection]\label{thm:triadic}
$\sigma^2$ on $\Omega$ has fixed-point set $\{0,3,8,9\}$ and $3$-cycles
$(1\,6\,4)$, $(7\,5\,2)$. The orbit of $\mathrm{HARMONY}=7$ is
$\{7,5,2\}$.
\end{theorem}

\begin{proof}
Direct computation from $\sigma=(0)(3)(8)(9)(1\,7\,6\,5\,4\,2)$
(Theorem~\ref{thm:cs}); squaring the $6$-cycle gives the two
$3$-cycles. The Conservation Tetrad is fixed by both $\sigma$ and
$\sigma^2$.
\end{proof}

\begin{theorem}[$4$-core bridge identity]\label{thm:bridge}
The Conservation Tetrad $\{0,3,8,9\}$ and the joint TSML+BHML $4$-core
$\{0,7,8,9\}$ of \cite{SandersGishFourCore} differ exactly by the
$\mathrm{PROGRESS}/\mathrm{HARMONY}$ swap:
\[
\{0,3,8,9\} \;\triangle\; \{0,7,8,9\} \;=\; \{3,7\}.
\]
\end{theorem}

\begin{proof}
Set-theoretic. The two sets share $\{0,8,9\}$; the unique
disagreement is the $\sigma^2$-fixed element $3$ (in the
Conservation Tetrad) versus the BHML/TSML-jointly-absorbed element
$7$ (in the $4$-core).
\end{proof}

%==============================================================
\section{Orphan~6: The $71$-cell field WOBBLE}
\label{sec:wobble}
%==============================================================

\begin{definition}[Field WOBBLE]\label{def:wobble}
The \emph{field WOBBLE} between $\mathrm{CL}_\TSML$ and
$\mathrm{CL}_\BHML$ is the symmetric-difference cell-count
\[
\mathrm{WOBBLE}(\TSML, \BHML)
\;:=\; \bigl|\{(i, j) \in \Omega^2 : \mathrm{CL}_\TSML[i,j]
\neq \mathrm{CL}_\BHML[i,j]\}\bigr|.
\]
\end{definition}

\begin{theorem}[$\mathrm{WOBBLE} = 71$]\label{thm:wobble}
$\mathrm{WOBBLE}(\TSML, \BHML) = 71$. Equivalently, the two tables
agree on exactly $29$ cells and disagree on the remaining $71$.
\end{theorem}

\begin{proof}
Direct cell-by-cell comparison of $\mathrm{CL}_\TSML$ (\cite[\S2]{SandersForcing})
and $\mathrm{CL}_\BHML$ (\cite[\S2.2]{SandersGishSO10}) on all $100$
cells. The dependence on the axioms is via A1--A8 for both tables;
A9 distinguishes the BUMP-position structure but does not enter the
cell-count itself.
\end{proof}

\begin{remark}[The DOING-rate near-coincidence]
\label{rem:doing-rate}
Restrict to the off-absorbing, off-diagonal cells of the substrate
(``DOING-cells''). Their count depends on whether one includes or
excludes certain boundary cells; for one natural normalization the
cell-count $N$ satisfies $71/N$ within $\pm 1\%$ of $T^* = 5/7
\approx 0.7143$. We do \emph{not} assert any algebraic identity:
the empirical observation is $71/100 = 0.71$, near to but not equal
to $5/7$. Settling whether $71/N = 5/7$ exactly for some
structurally-natural $N$ is open. Earlier drafts of this paper
named a Theorem~6.3 around this near-coincidence; it is correctly
stated as a Remark.
\end{remark}

%==============================================================
\section{Discussion}\label{sec:disc}
%==============================================================

\subsection{Why bundle?}
Each of the six structural facts is short. Individually each is too
brief for a standalone publication. Bundled, they form the structural
fingerprint of the bimodal $\alpha_A$ gap of
Conjecture~\ref{conj:bimodal} and a citable reference for the
downstream J-series \cite{SandersGishSO8,SandersGishSO10,SandersGishFourCore}.

\subsection{What this bundle does \emph{not} do}
We do not prove Conjecture~\ref{conj:bimodal}. The bundle records the
structural facts that any future proof will need to use; it does not
itself give a proof of structural exclusion in the band
$(0.5,\,0.87)$. The natural next paper is the
bimodal-$\alpha_A$-gap conjecture as a standalone foundational paper
\cite{FamilyStructureV1}.

\subsection{Connection to the parallel-substrate / lens architecture}
Three of the orphans (\S\ref{sec:std}, \S\ref{sec:ladder},
\S\ref{sec:wobble}) directly involve the three-substrate structure:
\S\ref{sec:std} names the third table; \S\ref{sec:ladder} carries
both TSML and BHML rungs (and Galois content from the LMFDB
quartic); \S\ref{sec:wobble} quantifies the TSML/BHML disagreement.
Two orphans (\S\ref{sec:cycle-skeleton}, \S\ref{sec:triadic}) are
$\TSML$-only structural facts. \S\ref{sec:bdc} is a $\STD$-only
fact.

\subsection{Future work}
Two further orphans -- the PathPair / LensTrace API, and the
joint eight-shell sub-magma chain -- are deferred to companion
papers because they require runtime infrastructure or exhaustive
sub-magma enumeration beyond the scope of a short note. The
eight-shell chain (sizes $\{1,4,5,6,7,8,9,10\}$ with forbidden sizes
only $\{2,3\}$) appears in \cite{SandersGishFourCore}; the
PathPair / LensTrace API is the subject of a separate paper in
preparation.

%==============================================================
\section{Reproducibility}\label{sec:repro}
%==============================================================

All six structural facts are verified in
\texttt{Gen13/targets/foundations/invariants.py}
\cite{FoundationsModule}. The full $48$-invariant verification suite
runs in under $30$ seconds on a standard laptop; it reports all $48$
invariants passing as of 2026-05-06, and includes the six
orphan-corresponding invariants (the three-substrate triple, the
$71$-roles signature, the $\sigma$-orbit decompositions, the BDC
numerical constants, the $\sigma^2$-triadic structure, and the
WOBBLE count).

%==============================================================
\section*{Acknowledgments}
%==============================================================
The structural framing of the (TSML, BHML) family as a five-criterion
membership space with the $4$-core at $\alpha_M = \tfrac12$ as its
center, and the identification of the bimodal $\alpha_A$ gap as an
open structural question, owes substantially to discussions with an
external collaborator in May 2026. The structural facts collected
here have accumulated over multiple internal sprints (Sprints 14--18)
and through the lens-taxonomy session of May 2026; we thank the
collaborators of those sprints for the structural verifications that
produced each fact.

%==============================================================
\begin{thebibliography}{99}
%==============================================================

\bibitem{SandersForcing}
B.~R.~Sanders and M.~Gish,
\emph{The CL Forcing Axioms: A1--A9 Uniquely Force the Canonical Composition Lattice},
Submitted to \emph{Algebraic Combinatorics} (2026)
[J25 of the J-series].

\bibitem{SandersGishSO8}
B.~R.~Sanders and M.~Gish,
\emph{$\mathfrak{so}(8) = D_4$ from the Antisymmetrized Closure of a Canonical $\Z/10\Z$ Magma},
Submitted to \emph{Journal of Algebra} (2026)
[J29 of the J-series].

\bibitem{SandersGishSO10}
B.~R.~Sanders and M.~Gish,
\emph{Joint Lie Closure of a Pair of $\Z/10\Z$ Magmas: an $\mathfrak{so}(10)$ Identification},
Submitted to \emph{Israel Journal of Mathematics} (2026)
[J30 of the J-series].

\bibitem{SandersGishFourCore}
B.~R.~Sanders and M.~Gish,
\emph{The $4$-Core $\{0,7,8,9\}$: Joint TSML+BHML Closure and the Universal Attractor},
Submitted to \emph{Algebraic Combinatorics} (2026)
[J35/four-core paper of the J-series].

\bibitem{SandersGishFpExtension}
B.~R.~Sanders and M.~Gish,
\emph{$\mathbb{F}_p$ Extensions of $\mathrm{CL}_\BHML$: Universality Across Six Prime Fields},
Submitted to \emph{Communications in Algebra} (2026)
[J34 of the J-series].

\bibitem{FamilyStructureV1}
\emph{TIG Family Structure: Membership, Center, Boundaries (v1)},
Internal reference document
\texttt{Atlas/META\_PLAN\_2026-05-06/FAMILY\_STRUCTURE\_v1.md},
adopted as foundational structural reference 2026-05-07.

\bibitem{FoundationsModule}
\emph{Foundations 48-Invariant Module},
\texttt{Gen13/targets/foundations/invariants.py} (and
\texttt{cl\_std.py} for $\mathrm{CL}_\STD$ data); CK research
codebase, 2026-05-06.

\bibitem{DrapalWanless2021}
A.~Dr\'apal and I.~M.~Wanless,
\emph{Maximally non-associative quasigroups},
J. Combinatorial Theory, Series A \textbf{184} (2021), 105510.

\bibitem{LMFDB}
The LMFDB Collaboration,
\emph{The L-functions and Modular Forms Database, number field 4.2.10224.1},
\url{https://www.lmfdb.org/NumberField/4.2.10224.1}, 2024.

\bibitem{SandersForcingOld}
Earlier internal draft of the present paper (2026-04 working version),
preserved in the J28 manuscript folder; superseded by the present
revision.

\end{thebibliography}

\end{document}
