%%% ============================================================
%%% The Sinc^2 Zero Law in Prime Arithmetic
%%% Brayden Ross Sanders (7Site LLC, Hot Springs AR), 2026
%%% DOI: 10.5281/zenodo.18852047
%%% Target venue: Integers -- Electronic Journal of
%%%               Combinatorial Number Theory
%%% Source: WP_SINC2_ZERO_LAW.md (Gen12/journal_attempts/01)
%%% Verification: proof_d25_loop_closure.py (primes 3..199)
%%% MSC: 11A41, 11N05, 42A16
%%%
%%% === STATUS: PULLED BACK FROM 2026-04-22 SUBMISSION QUEUE ===
%%% Reason (2026-04-19 pre-push audit): the central biconditional
%%%   sinc^2(k/n)=0 <=> n | k
%%% holds for EVERY positive integer n, not only for primes, so the
%%% main theorem is trivially true and the "prime" qualifier is
%%% vacuous. The sinc^2 zeros do not distinguish primes from
%%% composites; the prime-specific content must come from a First-G
%%% event formulation that is not yet drafted in this manuscript.
%%%
%%% ACTION: this manuscript is retained as a working draft; a
%%% rewrite to a non-trivial prime-specific statement (the First-G
%%% event replacement theorem sketched in
%%% Atlas/SINC2_SHARPEN_DECISION_2026_04_19.md sec.4) is required
%%% before any Integers submission. Do NOT submit in current form.
%%% === END PULL-BACK NOTICE ===
%%% ============================================================

¥documentclass[11pt,reqno]{amsart}

¥usepackage{amsmath,amssymb,amsthm,mathtools}
¥usepackage[margin=1in]{geometry}
¥usepackage[colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
¥usepackage{microtype}

%% -------- theorem environments --------
¥theoremstyle{plain}
¥newtheorem{theorem}{Theorem}
¥newtheorem{corollary}[theorem]{Corollary}
¥newtheorem{lemma}[theorem]{Lemma}

¥theoremstyle{remark}
¥newtheorem*{remark}{Remark}

%% -------- macros --------
¥DeclareMathOperator{¥sinc}{sinc}
¥newcommand{¥Z}{¥mathbb{Z}}
¥newcommand{¥N}{¥mathbb{N}}
¥newcommand{¥Q}{¥mathbb{Q}}

%% -------- title matter --------
¥title[The sinc¥texorpdfstring{$^2$}{¥^{}2} zero law]{The ¥texorpdfstring{$¥sinc^2$}{sinc¥^{}2} Zero Law in Prime Arithmetic}

¥author{Brayden Ross Sanders}
¥address{7Site LLC, Hot Springs, Arkansas, USA}
¥email{brayden@7site.co}

¥date{¥today}

¥keywords{sinc function, prime arithmetic, divisibility,
Fourier coefficients}

¥subjclass[2020]{11A41, 11N05, 42A16}

¥begin{document}

¥begin{abstract}
We prove that for any prime $p$ and any integer $k$, the squared
sinc function satisfies
¥[
  ¥sinc^2(k/p)=0 ¥iff p¥mid k.
¥]
Within the corridor $k¥in¥{1,¥dots,p¥}$ the unique zero occurs at
the endpoint $k=p$; every interior position $k<p$ is provably
nonzero --- a consequence of primality, not observation. We derive
three corollaries: ¥emph{loop closure} (the corridor closes exactly
once, at the prime), ¥emph{fold necessity} (a unique amplitude
crossing occurs in the interior of every prime corridor), and the
¥emph{no-shortcut lemma} (the path from the corridor entrance to its
zero has length exactly $p-1$). The main theorem is verified by
exact arithmetic for all primes $p¥in¥{3,5,7,¥dots,199¥}$ with zero
exceptions; the verification script
¥texttt{proof¥_d25¥_loop¥_closure.py} is supplied as electronic
supplementary material.
¥end{abstract}

¥maketitle

%% -------- section 1: setup --------
¥section{Setup}¥label{sec:setup}

The normalized sinc function is defined as
¥begin{equation}¥label{eq:sinc-def}
  ¥sinc(x)=
  ¥begin{cases}
     ¥dfrac{¥sin(¥pi x)}{¥pi x} & x¥neq 0,¥¥[4pt]
     1 & x=0,
  ¥end{cases}
¥end{equation}
and its square $¥sinc^2(x)=¥sinc(x)^2$. The function appears in
signal processing as the Shannon-sampling kernel
¥cite{Shannon1949}, in random matrix theory as the complement of
Montgomery's pair-correlation function for Riemann zeros
¥cite{Montgomery1973}, and --- as shown here --- directly in prime
arithmetic.

For a prime $p$ and integer $k¥ge 1$ we study the values
$¥sinc^2(k/p)$ along the rational corridor
$¥{1/p,¥,2/p,¥,¥dots¥}$.

%% -------- section 2: main theorem --------
¥section{The Main Theorem}¥label{sec:main}

¥begin{theorem}[Sinc$^2$ Zero Law]¥label{thm:main}
Let $p$ be prime and $k$ a positive integer. Then
¥begin{equation}¥label{eq:main}
  ¥sinc^2(k/p)=0 ¥iff p¥mid k.
¥end{equation}
Within $k¥in¥{1,¥dots,p¥}$ the unique zero is at $k=p$.
¥end{theorem}

¥begin{proof}
Since $k/p¥neq 0$ for $k¥ge 1$, we have
¥[
  ¥sinc^2(k/p)=¥frac{¥sin^2(¥pi k/p)}{(¥pi k/p)^2}.
¥]
This vanishes if and only if $¥sin(¥pi k/p)=0$, equivalently
$¥pi k/p¥in¥pi¥Z$, equivalently $k/p¥in¥Z$, equivalently
$p¥mid k$.

For $k¥in¥{1,¥dots,p-1¥}$: since $p$ is prime and $1¥le k<p$, we
have $¥gcd(k,p)=1$, so $p¥nmid k$, so $¥sinc^2(k/p)>0$.

At $k=p$: $k/p=1¥in¥Z$, so $¥sinc^2(1)=0$.
¥end{proof}

¥begin{remark}
The argument uses primality at exactly one step: $¥gcd(k,p)=1$ for
$k<p$. This fails for composite moduli. If $n=pq$ and $k=p<n$, then
$¥gcd(k,n)=p>1$, but $k/n=p/(pq)=1/q¥notin¥Z$, so
$¥sinc^2(k/n)>0$. The $¥sinc^2$ zero law identifies
¥emph{divisibility}, not just coprimality. What primality
contributes is that no proper divisor of $p$ lies strictly between
$1$ and $p$, so the corridor's interior is clean.
¥end{remark}

%% -------- section 3: corollaries --------
¥section{Corollaries}¥label{sec:cor}

¥begin{corollary}[Loop Closure]¥label{cor:loop}
The corridor $¥{1/p,2/p,¥dots,p/p¥}$ is nonzero on
$¥{1/p,¥dots,(p-1)/p¥}$ and zero at $p/p=1$. The corridor closes
exactly once.
¥end{corollary}

¥begin{proof}
Immediate from Theorem‾¥ref{thm:main}: the corridor $k/p$ as $k$
runs from $1$ to $p$ begins in positive territory and terminates at
zero, giving one crossing at the prime itself.
¥end{proof}

¥begin{corollary}[Fold Necessity]¥label{cor:fold}
For every prime $p¥ge 3$, there exists a unique $x^{*}¥in(0,1)$ ---
the ¥emph{fold} --- where $¥sinc^2(x^{*})=1/2$. This fold lies in the
interior of the corridor: there exist $k_1,k_2<p$ with
$¥sinc^2(k_1/p)>1/2>¥sinc^2(k_2/p)$.
¥end{corollary}

¥begin{proof}
$¥sinc^2$ is continuous and strictly decreasing on $(0,1)$: the
function $¥sin(¥pi x)$ dominates $¥pi x$ near $0$ and they cross at
$x=1$. Since $¥sinc^2(0)=1>1/2$ and $¥sinc^2(1)=0<1/2$, the
intermediate value theorem gives a unique $x^{*}¥in(0,1)$ with
$¥sinc^2(x^{*})=1/2$.

We note that $x^{*}¥neq 1/2$: direct computation gives
¥[
  ¥sinc^2(1/2)
    =¥left(¥frac{¥sin(¥pi/2)}{¥pi/2}¥right)^{¥!2}
    =¥left(¥frac{2}{¥pi}¥right)^{¥!2}
    =¥frac{4}{¥pi^2}¥approx 0.4053,
¥]
which is strictly less than $1/2$. For $p=7$ one checks
$¥sinc^2(3/7)¥approx 0.524>1/2$ and
$¥sinc^2(4/7)¥approx 0.295<1/2$, so the fold crosses between
$k=3$ and $k=4$.
¥end{proof}

¥begin{corollary}[No interior zero]¥label{cor:noshortcut}
For every prime $p$ and every integer $k$ with $1¥le k¥le p-1$,
$¥sinc^2(k/p)¥neq 0$.
¥end{corollary}

¥begin{proof}
Immediate from Theorem‾¥ref{thm:main}: the biconditional
$¥sinc^2(k/p)=0¥iff p¥mid k$ rules out any interior $k$ with
$1¥le k¥le p-1$, since $p¥nmid k$ for $k$ in that range.
¥end{proof}

%% -------- section 4: boundary value --------
¥section{The Boundary Value ¥texorpdfstring{$¥sinc^2(1/2)=4/¥pi^2$}{sinc¥^{}2(1/2) = 4/pi¥^{}2}}¥label{sec:boundary}

The fold amplitude $¥sinc^2(1/2)$ has a closed form:
¥begin{equation}¥label{eq:fourpi2}
  ¥sinc^2(1/2)
    =¥left(¥frac{¥sin(¥pi/2)}{¥pi/2}¥right)^{¥!2}
    =¥left(¥frac{2}{¥pi}¥right)^{¥!2}
    =¥frac{4}{¥pi^2}¥approx 0.4053.
¥end{equation}
This is a transcendental number. It is the amplitude of the first
sidelobe of a rectangular spectral gate --- a classical fact from
signal processing --- but it appears here as the natural boundary
value dividing corridor positions that exceed $1/2$ from those that
do not.

Montgomery's pair-correlation function for Riemann zeros is
$R_2(u)=1-¥sinc^2(u)$ ¥cite{Montgomery1973}. The complement
$¥sinc^2(u)$ and $R_2(u)$ sum to unity --- a complete spectral
partition. The boundary $4/¥pi^2=¥sinc^2(1/2)$ appears in both
frameworks, derived rather than imposed.

%% -------- section 5: First-G Law --------
¥section{Connection to the First-G Law}¥label{sec:firstG}

The First-G Law ¥cite{SandersWP34} establishes an equivalent fact
in coprimality-partition language: for any semiprime $b=pq$ with
$p¥le q$, the first non-unit element in $¥{1,¥dots,k¥}$ with
respect to $b$ appears at exactly $k=p$. The $¥sinc^2$ zero law is
the continuum-limit statement of the same algebraic fact. In the
limit $p¥to¥infty$, the harmonic pre-echo function
¥begin{equation}¥label{eq:R}
  R(k,p)¥;=¥;¥frac{¥sin^2(¥pi k/p)}{k^2¥,¥sin^2(¥pi/p)}
  ¥;¥longrightarrow¥; ¥sinc^2(k/p)
¥end{equation}
recovers the $¥sinc^2$ field exactly ¥cite{SandersWP35}. The two
results are two faces of the same structure --- one discrete, one
continuous.

%% -------- section 6: verification --------
¥section{Verification}¥label{sec:verify}

Theorem‾¥ref{thm:main} is verified for all primes
$p¥in¥{3,5,7,11,¥dots,199¥}$ ($46$ primes) by exact arithmetic in
Python's ¥texttt{fractions.Fraction} and ¥texttt{sympy} modules.
For each prime‾$p$:
¥begin{itemize}¥setlength¥itemsep{1pt}
  ¥item All $k¥in¥{1,¥dots,p-1¥}$: $¥sinc^2(k/p)>0$ confirmed by
        exact computation.
  ¥item $k=p$: $¥sinc^2(1)=0$ confirmed.
  ¥item Zero exceptions across all $46$ primes.
¥end{itemize}

The runnable verification
(¥texttt{proof¥_d25¥_loop¥_closure.py}, approximately 270 lines)
is supplied as electronic supplementary material and is also
archived at DOI‾¥texttt{10.5281/zenodo.18852047}.

%% -------- section 7: limits --------
¥section{What This Result Does Not Claim}¥label{sec:limits}

This paper does not claim: a new proof of the infinitude of
primes; a connection to the Riemann Hypothesis beyond the
Montgomery bridge observation in Section‾¥ref{sec:boundary}; that
the fold position $x^{*}$ of Corollary‾¥ref{cor:fold} is
computable in closed form; or any result about the distribution
of primes. Theorem‾¥ref{thm:main} is a finite algebraic fact about
$¥sinc^2$ evaluated at rational points with prime denominator.

%% -------- references --------
¥begin{thebibliography}{99}

¥bibitem{Montgomery1973}
H.‾L.‾Montgomery.
¥newblock The pair correlation of zeros of the zeta function.
¥newblock In ¥emph{Analytic Number Theory}, Proceedings of Symposia
in Pure Mathematics, volume‾24, pages 181--193. American
Mathematical Society, Providence, RI, 1973.

¥bibitem{SandersWP34}
B.‾R.‾Sanders.
¥newblock WP34 --- The First-G Law and Prime-Forced Dispersion.
¥newblock 7Site Research Collaboration, 2026.
¥newblock DOI: ¥texttt{10.5281/zenodo.18852047}.

¥bibitem{SandersWP35}
B.‾R.‾Sanders, C.‾A.‾Luther, and M.‾Gish.
¥newblock WP35 --- The Prime Phase Transition: Harmonic Pre-Echo,
Zero-Width Gates, and the Geometry of RSA Security.
¥newblock 7Site Research Collaboration, 2026.
¥newblock DOI: ¥texttt{10.5281/zenodo.18852047}.

¥bibitem{Shannon1949}
C.‾E.‾Shannon.
¥newblock Communication in the presence of noise.
¥newblock ¥emph{Proceedings of the IRE}, 37(1):10--21, 1949.

¥end{thebibliography}

%% -------- end matter --------
¥vfill
¥noindent
¥footnotesize{%
¥textsc{Provenance.}¥; Source manuscript
¥texttt{WP¥_SINC2¥_ZERO¥_LAW.md} in the 7Site Research repository
(DOI‾¥texttt{10.5281/zenodo.18852047}). External citations are
drawn from ¥texttt{Atlas/ATLAS¥_CITATIONS.md} (¥S A analytic number
theory; Shannon, Montgomery). Internal cross-references
(sinc$^2$ zero law, prime corridor closure, fold necessity,
no-shortcut lemma) carry master-register numbering per
¥texttt{Atlas/MASTER¥_ATLAS¥_v3¥_5¥_2026¥_04¥_18.md}. Readiness
flag‾[fire -- submit-ready], Sprint‾34 ¥textquotedblleft Ship the
First Three.¥textquotedblright}

¥end{document}
