% J17 — Total-Dimension Match V^{⊗n} and Cl(2n), with Refined-Cell Grading
%
% Brayden R. Sanders, M. Gish, 7Site LLC
% Submitted to: Linear Algebra and its Applications
% Source corpus: WP119 (Gen12/targets/clay/papers/sprint18_bridge_dirac_2026_05_04/journals/WP119_journal_clean.tex)
% Verifier: tig_dirac.tensor_partition() at Gen13/targets/ck/brain/dirac/tig_dirac.py
% Verification: test_tig_dirac.py tests T13, T14, T15
% Companion cited as already-submitted: J23
% MSC 2020: 15A66, 17A30, 11T55
% R1 revision 2026-05-07: corrected Theorem 3.2 (binomial-vs-grade), restructured §3-§4
% per fresh-eyes referee report J17_LinAlgApps_FreshEyes.md.

\documentclass[11pt,reqno]{amsart}

\usepackage{amsmath, amssymb, amsthm, mathtools}
\usepackage[margin=1in]{geometry}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{microtype}
\usepackage{booktabs}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}

\theoremstyle{remark}
\newtheorem*{remark}{Remark}

\newcommand{\Z}{\mathbb{Z}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Cl}{\mathrm{Cl}}

\title[Total-dimension match $V^{\otimes n}$ and $\Cl(2n)$]
{Total-Dimension Match Between Tensor Powers of a \\
 Finite-Field 4-Algebra and Real Clifford Algebras $\Cl(2n)$, \\
 with a Refined-Cell Grading}

\author{Brayden R.\ Sanders \and M.\ Gish}
\address{7Site LLC, Hot Springs, Arkansas, USA}
\email{brayden@7site.co}

\author{Brayden R.\ Sanders \and M.\ Gish}
\address{Independent Researcher, Hot Springs, Arkansas, USA}
\email{monica.gish1992@gmail.com}

\subjclass[2020]{15A66, 17A30, 11T55}
\keywords{Clifford algebra, tensor power, finite-field algebra,
binomial decomposition, idempotent decomposition}

\date{\today}

\begin{document}

\begin{abstract}
We record an arithmetic identity and a refined-cell construction
relating tensor powers $V^{\otimes n}$ of a particular 4-dimensional
commutative non-associative algebra $V$ over $\F_5$ to real Clifford
algebras $\Cl(2n, 0)$.
At the level of total dimension,
\[
\dim_{\F_5} V^{\otimes n} \;=\; 4^n \;=\; 2^{2n} \;=\; \dim_\R \Cl(2n)
\quad\text{for every } n \geq 0,
\]
which is forced by $\dim V = 4 = 2^2$.
The non-trivial content is a \emph{refined cell decomposition} of
$V^{\otimes n}$: each tensor slot carries two structural sign bits
(membership of the two orthogonal idempotents and the two nilpotent
basis elements), giving $4^n = 2^{2n}$ one-dimensional refined cells
whose Hamming-weight distribution by the $2n$ structural bits is
exactly the binomial sequence $\binom{2n}{k}$, $k = 0, \ldots, 2n$,
matching the grade dimensions of $\Cl(2n)$.

We are explicit about what is not proved.
The map sending refined cells of $V^{\otimes n}$ to $\Cl(2n)$ basis
multivectors is, at present, only a bookkeeping bijection on
basis-element labels: it is well-defined, but we do not prove that it
extends to a structure-preserving map of $\F_5$-vector spaces or
algebras to a Clifford algebra over a common field.
The match between coarse-cell weights $\binom{n}{k}$ and the SU(5)
representation dimensions at $n=5$ is a binomial-coefficient
coincidence (Remark \ref{rem:su5}), not a representation-theoretic
theorem.

The construction is verified at $n = 0, 1, 2, 3, 4, 5$ in the
accompanying script \texttt{test\_tig\_dirac.py} and via the reference
implementation \texttt{tig\_dirac.tensor\_partition()} and the new
\texttt{refined\_cell\_distribution()} helper.
\end{abstract}

\maketitle

\tableofcontents

%==============================================================
\section{Introduction}
%==============================================================

The real Clifford algebras $\Cl(p, q)$ over a quadratic form of
signature $(p, q)$ form a periodic family under
$\Cl(p, q) \otimes \Cl(0, 0) \cong \Cl(p, q)$ and Bott periodicity
$\Cl(n + 8) \cong \Cl(n) \otimes \Cl(8)$, with applications spanning
geometric algebra (Hestenes--Sobczyk \cite{HesteneSobczyk1984}),
spinor representations of orthogonal groups, quantum-information
stabilizer codes, and Standard-Model gauge structure.

This paper records an elementary structural observation about a
particular 4-dimensional commutative non-associative algebra $V$ over
$\F_5$ studied in the companion paper \cite{SandersGishDiscreteDirac}.
The total dimension of every tensor power $V^{\otimes n}$ matches
$\dim_\R \Cl(2n)$, and a natural refinement of the basis-cell
decomposition of $V^{\otimes n}$ into $4^n = 2^{2n}$ one-dimensional
cells carries a Hamming-weight grading whose dimension distribution
$\binom{2n}{k}$ matches the grade dimensions of $\Cl(2n)$.

The paper is short, and the results are elementary in the technical
sense.  We have nonetheless found the structural picture useful in
companion work \cite{SandersGishDiscreteDirac, SandersGishFpUniversality},
where the algebra $V$ is the substrate of a discrete Dirac construction
on the 4-core of $\Z/10\Z$.  The present note isolates the
linear-algebraic content (Theorems \ref{thm:total-dim} and
\ref{thm:refined}) so that it can be checked or used independently of
the broader framework.

\subsection*{Scope statement.}
We do \emph{not} construct a structure-preserving map
$V^{\otimes n} \to \Cl(2n)$ in this paper.  The relationship
established below is a basis-cell bijection equipped with a Hamming
grading, and an arithmetic identity at the level of total dimension.
A genuine algebra map between $V^{\otimes n}$ over $\F_5$ and
$\Cl(2n)$ over $\R$ would have to address the change of base ring;
this is left for future work.  See \S\ref{sec:open}.

\subsection*{Notation.}
$V$ denotes the 4-dimensional commutative non-associative algebra over
$\F_5$ studied in \cite{SandersGishDiscreteDirac}; the multiplication
is given by an explicit $4 \times 4$ table on the basis
$\{e_0, e_2, e_3, e_4\}$ recalled in \S\ref{sec:V}.
$\Cl(2n)$ always means $\Cl(2n, 0)$ over $\R$ -- $2n$ generators each
squaring to $+1$.

%==============================================================
\section{The algebra $V$ and its idempotent decomposition}
\label{sec:V}
%==============================================================

The 4-dimensional algebra $V$ over $\F_5$ has 4 idempotents (one zero,
three non-zero), a 1-dim Grassmann annihilator $\F_5 \cdot e_0$, and a
1-dim associator image; the multiplication table is given in
\cite[Table~1]{SandersGishDiscreteDirac}.  For the present paper,
the only data we need are the following two direct-sum decompositions.

\paragraph{Coarse decomposition (2 summands).}
\[
V \;=\; \underbrace{(\F_5\, e_2 \oplus \F_5\, e_3)}_{V_+}
        \;\oplus\;
        \underbrace{(\F_5\, e_0 \oplus \F_5\, e_4)}_{V_-},
\]
where $V_+$ is the 2-dimensional subspace containing the primitive
idempotent $e_2$ and the nilpotent $e_3$, and $V_-$ is its complement.
This is the decomposition used in
\cite[\S2]{SandersGishDiscreteDirac}; we record it without claiming
it is canonical (see Remark \ref{rem:choice} below).

\paragraph{Refined decomposition (4 summands).}
\[
V \;=\; \F_5\, e_0 \;\oplus\; \F_5\, e_2 \;\oplus\; \F_5\, e_3 \;\oplus\; \F_5\, e_4,
\]
the $\F_5$-vector-space splitting into the four basis lines.

The coarse decomposition gives, across $n$ tensor factors, a
\emph{coarse cell partition} of $V^{\otimes n}$ into $2^n$ subspaces,
each of dimension $2^n$ (label each tensor slot by a sign
$\sigma_i \in \{+, -\}$ recording $V_+$ or $V_-$).
The refined decomposition gives a \emph{refined cell partition} into
$4^n$ one-dimensional subspaces (label each slot by a 2-bit string
$(a_i, b_i) \in \{0,1\}^2$ recording which of the four basis lines).

We refer to the $2^n$ coarse cells as \emph{coarse cells} and the
$4^n$ refined cells as \emph{refined cells}.  In the previous draft of
this paper the coarse cells were called \emph{fine cells}; the change
in nomenclature is forced by Theorem \ref{thm:refined} below, which
shows the \emph{refined} cells are the right object for the
Clifford comparison.

\begin{remark}\label{rem:choice}
Both decompositions are basis-dependent.  The refined decomposition is
the canonical splitting of $V$ into $\F_5$-lines determined by the
basis $\{e_0, e_2, e_3, e_4\}$.  The coarse decomposition is one of
$\binom{4}{2}/2 = 3$ ways to partition the four basis lines into two
pairs; the pairing $\{e_2, e_3\} \,|\, \{e_0, e_4\}$ used here is the
one inherited from \cite{SandersGishDiscreteDirac}, in which $e_2$
(the privileged primitive idempotent) and $e_3$ (its companion in the
2-dim ``particle'' subalgebra) are grouped together.
The choice does not affect total dimensions.
\end{remark}

%==============================================================
\section{Total dimension match}
%==============================================================

\begin{theorem}[Total-dimension match]
\label{thm:total-dim}
For every $n \geq 0$,
\[
\dim_{\F_5} V^{\otimes n} \;=\; 4^n \;=\; 2^{2n} \;=\; \dim_\R \Cl(2n).
\]
\end{theorem}

\begin{proof}
$\dim_{\F_5} V^{\otimes n} = (\dim_{\F_5} V)^n = 4^n$ by tensor-product
dimension. $4^n = (2^2)^n = 2^{2n}$. $\dim_\R \Cl(2n) = 2^{2n}$ by the
classical formula for the dimension of a Clifford algebra over a
quadratic form on $2n$ generators
\cite[Ch.~I]{HesteneSobczyk1984}.
\end{proof}

\begin{remark}
Theorem~\ref{thm:total-dim} is forced by $\dim V = 4 = 2^2$.  The same
identity holds with $V$ replaced by any 4-dimensional algebra over any
field.  The \emph{refinement} below is what depends on the specific
structure of $V$ -- in particular, on its having an
$\F_5$-vector-space splitting into four lines.
\end{remark}

%==============================================================
\section{Refined-cell grading matches $\Cl(2n)$ grade dimensions}
%==============================================================

\begin{theorem}[Refined-cell binomial grading]
\label{thm:refined}
For every $n \geq 0$ the $4^n = 2^{2n}$ refined cells of
$V^{\otimes n}$ partition into Hamming-weight classes whose
multiplicities match the grade dimensions of $\Cl(2n)$:
\[
\#\{\text{refined cells of weight } k\} \;=\; \binom{2n}{k}
\;=\; \dim_\R \Cl^{(k)}(2n), \qquad k = 0, 1, \ldots, 2n.
\]
Here the \emph{weight} of a refined cell labeled by $(a_1, b_1, \ldots,
a_n, b_n) \in \{0,1\}^{2n}$ is $a_1 + b_1 + \cdots + a_n + b_n$.
\end{theorem}

\begin{proof}
A refined cell of $V^{\otimes n}$ is labeled by a pair $(a_i, b_i)$
for each tensor slot $i = 1, \ldots, n$, recording which of the four
basis lines $\F_5 e_0, \F_5 e_2, \F_5 e_3, \F_5 e_4$ that slot
contributes.  The total label is therefore a binary string of length
$2n$, and the number of refined cells of total Hamming weight $k$ is
the number of length-$2n$ binary strings of weight $k$, i.e.\
$\binom{2n}{k}$.
The Clifford grade $\Cl^{(k)}(2n)$ is, by definition, the
$\R$-span of products of $k$ distinct generators among the $2n$
generators of $\Cl(2n)$, and has dimension $\binom{2n}{k}$
\cite[Ch.~I]{HesteneSobczyk1984}.
The two distributions are therefore equal as multisets of
non-negative integers indexed by $k = 0, \ldots, 2n$.
\end{proof}

\begin{remark}[On the coarse-cell distribution]
\label{rem:coarse}
The coarse-cell distribution of $V^{\otimes n}$ -- counting coarse
cells by the number of $V_+$-slots -- gives the binomial sequence
$\binom{n}{k}$ for $k = 0, \ldots, n$.  These are the multiplicities
of \emph{coarse} cells, each of which is itself a $2^n$-dimensional
subspace.  The $\binom{n}{k}$ sequence is \emph{not} the grade
sequence of $\Cl(2n)$, which is $\binom{2n}{k}$ over a $2n+1$-element
index set.  Theorem \ref{thm:refined} resolves the apparent mismatch
by working at the refinement: the coarse-cell distribution is the
``coarse shadow'' of the refined distribution, obtained by collapsing
the two bits per slot to one.  Concretely, the refined-cell weight $k$
splits as $k = k_+ + k_-$ where $k_+$ is the number of slots
contributing to $V_+$ (a coarse-cell coordinate) and $k_-$ is a
sub-grading inside that coarse cell; summing over $k_-$ recovers the
coarse multiplicity $\binom{n}{k_+}$ from the refined $\binom{2n}{k}$.
\end{remark}

\begin{remark}[Status of the bijection]\label{rem:bijection}
Theorem \ref{thm:refined} establishes equality of multiplicities
between two binomial distributions.  The natural bijection between
refined cells of $V^{\otimes n}$ and basis multivectors of $\Cl(2n)$
is by mapping the 2-bit slot-label $(a_i, b_i)$ to a pair of generators
of $\Cl(2n)$.  This bijection is well-defined on basis-element labels
but is not, as stated, a map of $\F_5$-vector spaces (the codomain is
an $\R$-vector space) and is not, as stated, an algebra homomorphism.
What it gives is a label-level identification of refined cells with
Clifford basis multivectors that respects the binomial grading.
A genuine algebra map between $V^{\otimes n}$ and a Clifford algebra
over a common field is left open; see \S\ref{sec:open}.
\end{remark}

%==============================================================
\section{The case $n = 5$}
\label{sec:n5}
%==============================================================

\begin{proposition}[Refined cells at $n = 5$]\label{prop:n5refined}
$V^{\otimes 5}$ has $4^5 = 1024$ refined cells whose
Hamming-weight distribution by the $10$ structural bits is
\[
\binom{10}{0},\binom{10}{1},\ldots,\binom{10}{10}
\;=\;
1, 10, 45, 120, 210, 252, 210, 120, 45, 10, 1.
\]
This matches the grade dimensions of $\Cl(10)$ exactly.
\end{proposition}

\begin{proof}
Specialization of Theorem \ref{thm:refined} to $n = 5$, $2n = 10$.
\end{proof}

\begin{proposition}[Coarse cells at $n = 5$]\label{prop:n5coarse}
$V^{\otimes 5}$ has $2^5 = 32$ coarse cells whose
$V_+$-weight distribution is
$\binom{5}{0}, \binom{5}{1}, \binom{5}{2}, \binom{5}{3},
 \binom{5}{4}, \binom{5}{5}
 = 1, 5, 10, 10, 5, 1$.
\end{proposition}

\begin{proof}
Coarse-cell version of Theorem \ref{thm:refined}; counts cells by the
number of $V_+$-slots.
\end{proof}

\begin{remark}[On the SU(5) coincidence]\label{rem:su5}
The numbers $1 + 5 + 10 + 10 + 5 + 1$ in
Proposition~\ref{prop:n5coarse} agree with the dimensions of the
SU(5) one-generation representation
$\mathbf{1} \oplus \bar{\mathbf{5}} \oplus \mathbf{10}$ together with
its conjugate.  This agreement is a binomial-coefficient identity:
$\binom{5}{0} = 1$, $\binom{5}{1} = 5$, $\binom{5}{2} = 10$, and
$\binom{5}{k} = \binom{5}{5-k}$.  It is true for any tensor structure
that admits a binary partition at each of $5$ tensor factors,
independent of $V$ or its multiplicative structure.  We record it as a
remark, not as a theorem: the present paper does not construct an
$\mathrm{SU}(5)$ action on $V^{\otimes 5}$, and we do not assert that
the binomial coincidence reflects any representation-theoretic content
of $V$.  Whether $V^{\otimes 5}$ admits a canonical $\mathrm{SU}(5)$
or $\mathrm{Spin}(10)$ action is open and is the subject of related
work \cite{SandersGishDiscreteDirac}.
\end{remark}

%==============================================================
\section{Verification}
%==============================================================

The accompanying scripts verify Theorems \ref{thm:total-dim} and
\ref{thm:refined} and Propositions \ref{prop:n5refined},
\ref{prop:n5coarse}:
\begin{itemize}
\item \texttt{test\_tig\_dirac.py} test T15: dimensional ladder for
$n = 0, 1, 2, 3, 4, 5$.
\item Test T13: $V^{\otimes n}$ has $2^n$ coarse cells for
$n = 1, \ldots, 5$.
\item Test T14: coarse-cell binomial $1 + 5 + 10 + 10 + 5 + 1 = 32$
at $n = 5$ (Proposition \ref{prop:n5coarse}).
\item Helper \texttt{tig\_dirac.refined\_cell\_distribution(n)}:
returns the refined-cell weight distribution
$\{k : \binom{2n}{k}\}$ for arbitrary $n$, used to verify
Theorem \ref{thm:refined} for $n = 0, \ldots, 5$.
\item \texttt{tig\_dirac.tensor\_partition()}: enumerates the
$2^n$ coarse cells.
\end{itemize}

All checks pass deterministically in $<1$ second on standard
hardware. The reference Python library \texttt{tig\_dirac.py} is
deposited at the public repository:
\url{https://github.com/TiredofSleep/ck/tree/tig-synthesis/Gen13/targets/ck/brain/dirac/tig_dirac.py}.

%==============================================================
\section{Open questions}
\label{sec:open}
%==============================================================

\begin{enumerate}
\item[(O1)]
\textbf{Structure-preserving map}.
Is there a natural map $V^{\otimes n} \otimes_{\F_5} K \to \Cl(2n; K)$,
for some choice of common ring $K$ (e.g.\ $\F_5$ replaced by an
algebraic closure of $\F_5$, or a Witt-vector lift to characteristic
$0$), under which the refined-cell decomposition of the left-hand side
maps to the grade decomposition of the right-hand side?
The bijection of basis labels in Theorem \ref{thm:refined} is the
candidate; what is missing is a verified product structure.

\item[(O2)]
\textbf{Action of $\mathrm{Spin}(2n)$}.
Does $V^{\otimes n}$ admit a canonical action of $\mathrm{Spin}(2n)$
or one of its $\F_5$-analogues, under which the coarse-cell or
refined-cell weight distributions correspond to weight spaces?
The case $n = 5$ would address Remark \ref{rem:su5}.

\item[(O3)]
\textbf{Periodicity}.
The 8-periodicity $\Cl(n+8) \cong \Cl(n) \otimes \Cl(8)$ on the
Clifford side has the trivial counterpart $V^{\otimes (n+8)} =
V^{\otimes n} \otimes V^{\otimes 8}$ on the algebra side.  Is there
content in the comparison beyond the trivial dimensional consequence
$4^{n+8} = 4^n \cdot 4^8$?
\end{enumerate}

%==============================================================
\section*{Acknowledgments}
%==============================================================

This work is part of a broader algebraic and physical framework
developed at 7Site LLC (Trinity Infinity Geometry, public repository
at github.com/TiredofSleep/ck). The present manuscript is
self-contained as a study in tensor algebras over finite fields;
familiarity with the broader framework is not required.

We thank an anonymous referee whose careful reading uncovered the
mis-statement in Theorem \ref{thm:refined} of an earlier draft (in
which the coarse-cell distribution $\binom{n}{k}$ was equated with the
$\Cl(2n)$ grade distribution $\binom{2n}{k}$); the present
refined-cell formulation directly addresses that mismatch.

%==============================================================
\begin{thebibliography}{99}
%==============================================================

\bibitem{HesteneSobczyk1984} D.~Hestenes, G.~Sobczyk,
\emph{Clifford Algebra to Geometric Calculus},
D.~Reidel Publishing Co., 1984.

\bibitem{GeorgiGlashow1974} H.~Georgi, S.~L.~Glashow,
\emph{Unity of all elementary-particle forces},
Phys.~Rev.~Lett.~\textbf{32} (1974), 438--441.

\bibitem{SandersGishDiscreteDirac} B.~R.~Sanders, M.~Gish,
\emph{Discrete Dirac on $\F_5^4$: Substrate Algebra of the 4-Core},
submitted to \emph{Algebras and Representation Theory}, 2026 (J23).

\bibitem{SandersGishFpUniversality} B.~R.~Sanders, M.~Gish,
\emph{F$_p$ Universality: The Operator-Substrate Construction over
Prime Fields}, submitted to \emph{Algebra Universalis}, 2026 (J21).

\bibitem{TIGsynthesis2026} B.~R.~Sanders,
\emph{Trinity Infinity Geometry Synthesis},
github.com/TiredofSleep/ck (default branch),
DOI: 10.5281/zenodo.18852047, 2026.

\end{thebibliography}

\end{document}
