% J33 — The CL Forcing Axioms: A1-A9 Uniquely Force the Canonical Composition Lattice
%
% Brayden R. Sanders, M. Gish, 7Site LLC
% Submitted to: Algebraic Combinatorics
% Source corpus: Atlas/LENS_TAXONOMY_2026-05-06/CL_FORCING_AXIOMS.md (2026-05-06)
% Companions cited as already-submitted: J09, J31
% MSC 2020: 08A40, 20N02, 17D99
% Tier: B (axiom-set forcing, Tier-A substrate elements explicitly flagged)

\documentclass[11pt,reqno]{amsart}

\usepackage{amsmath, amssymb, amsthm, mathtools}
\usepackage[margin=1in]{geometry}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{microtype}
\usepackage{booktabs}
\usepackage{array}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{conjecture}[theorem]{Conjecture}

\theoremstyle{remark}
\newtheorem*{remark}{Remark}

\newcommand{\Z}{\mathbb{Z}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\BUMP}{\mathrm{BUMP}}
\newcommand{\TSML}{\mathrm{TSML}}
\newcommand{\BHML}{\mathrm{BHML}}

\title[CL Forcing Axioms]
{The CL Forcing Axioms: A1--A9 Uniquely Force the Canonical Composition Lattice}

\author{Brayden R.~Sanders}
\address{7Site LLC, Hot Springs, Arkansas, USA}
\email{brayden@7site.co}

\author{M.~Gish}
\address{Independent Researcher, Hot Springs, Arkansas, USA}
\email{monica.gish1992@gmail.com}

\subjclass[2020]{08A40, 20N02, 17D99}
\keywords{composition lattice, finite magma, axiom forcing,
absorbing element, $\Z/10\Z$ substrate, BUMP enumeration}

\date{\today}

\begin{document}

\begin{abstract}
We isolate a finite list of nine axioms A1--A9 on a $10\times 10$
multiplication table over $\Z/10\Z$ and prove that they uniquely force
the canonical TSML composition lattice $\mathrm{CL}_{\TSML}$. The axioms
partition cleanly into two groups: \emph{absorbing-element rules}
(A2--A6, dictated by VOID's near-universal absorption with the
$(0,7)$ puncture and HARMONY's universal absorption) and
\emph{substrate-defining rules} (A7 diagonal HARMONY law, A8
HARMONY-default rule, A9 BUMP enumeration with explicit values at five
unordered positions $\{(1,2),(2,4),(2,9),(3,9),(4,8)\}$). The BUMP
\emph{positions} are forced by the BDC entropy-extremum condition
(Tier-B); the BUMP \emph{values} at those positions are
substrate-defining choices (Tier-A) that distinguish $\mathrm{CL}_\TSML$
from its parallel substrates $\mathrm{CL}_\BHML$ and $\mathrm{CL}_\mathrm{STD}$.
A cell-counting argument confirms A1--A9 fix all $100$ matrix entries.
This places $\mathrm{CL}_\TSML$ on the same axiomatic footing as
classical absorbing-element semigroups (\cite{HowieSemigroup}) while
making explicit the finite enumeration data needed to specify a
multiplication table on the punctured-VOID, HARMONY-saturated
$\Z/10\Z$ substrate.

The result feeds the three-substrate architecture of
\cite{SandersGishThreeSubstrate} (J31, Algebra Universalis) and the
lens-invariance framework of \cite{SandersLensInvariance}
(J09, JCT-A): the parallel substrates $\mathrm{CL}_\TSML$,
$\mathrm{CL}_\BHML$, $\mathrm{CL}_\mathrm{STD}$ all share A1--A6 but
diverge at A7--A9, giving a clean mechanism for the lens family.
\end{abstract}

\maketitle

\tableofcontents

%==============================================================
\section{Introduction}
%==============================================================

A composition lattice on a finite operator alphabet is a finite
multiplication table $M : \Omega \times \Omega \to \Omega$ that one
might hope to determine from a small list of axioms. When $\Omega$
contains a designated absorbing element (a ``zero'') and a designated
top element (a ``one''), the standard absorbing-element rules
\cite{HowieSemigroup, Clifford1961} fix the rows and columns of these
two elements but leave the rest of the table open.

In this paper we exhibit a small axiom set --- nine axioms total ---
that uniquely forces a specific $10 \times 10$ table on the substrate
$\Omega = \Z/10\Z$. The table $\mathrm{CL}_\TSML$ has the following
structural features:

\begin{itemize}
\item $73$ of $100$ cells take the value HARMONY ($=7$);
\item $10$ cells are \emph{BUMP cells}, where the value is non-HARMONY
and non-VOID;
\item the remaining $17$ cells are VOID, distributed along the row and
column of $0$ with one puncture at $(0,7)$.
\end{itemize}

The HARMONY-saturation rate $73/100$ is high but not maximal --- the
five unordered BUMP positions $\{(1,2), (2,4), (2,9), (3,9), (4,8)\}$
break it. These positions are not arbitrary: they coincide with the
information-theoretic extrema of the binary digit code (BDC) framework
applied to the substrate, where each BUMP cell carries
$\log_2(10/2) = 2.32$ bits of information beyond the HARMONY-default
baseline (and $3.50$ bits including the BDC structural entropy term).
This justifies the working slogan ``surprise IS information'':
the cells where the table fails to be HARMONY are exactly the cells
where it carries new information.

\subsection{What this paper proves.}
We give nine axioms A1--A9 (\S\ref{sec:axioms}) and show by direct
enumeration of $100$ cells that they have a unique solution
$\mathrm{CL}_\TSML$ as a $10\times 10$ matrix on $\Z/10\Z$
(\S\ref{sec:forcing}). Of those nine axioms, A1, A2, A3, A4, A7, and
the values in A9 are \emph{substrate-defining} (Tier-A: choose
differently and obtain a different substrate, e.g.\ $\mathrm{CL}_\BHML$
or $\mathrm{CL}_\mathrm{STD}$); A5, A6, A8, and the BUMP positions in
A9 are \emph{forced} (Tier-B: derived from A1--A4 plus commutativity
plus the BDC entropy extremum) (\S\ref{sec:tiering}).

\subsection{What this paper does \emph{not} prove.}
We do \emph{not} claim $\mathrm{CL}_\TSML$ is the unique
absorbing-element-saturated composition lattice on $\Z/10\Z$. Two
parallel substrates $\mathrm{CL}_\BHML$ (28 HARMONY) and
$\mathrm{CL}_\mathrm{STD}$ (44 HARMONY) satisfy A1, A2, A4 but
\emph{differ at} A7--A9. We discuss this three-substrate architecture
in \S\ref{sec:threetable} and refer to
\cite{SandersGishThreeSubstrate} for the full development. The lens
family of (TSML, BHML, STD)-style substrates is the natural object
forced by varying A7--A9-values while keeping A1--A6 fixed.

\subsection{Companion structure.}
This paper sits between two siblings:
\begin{itemize}
\item Backward: \cite{SandersLensInvariance} (J09, JCT-A) defines
\emph{lens invariance} for combinatorial statements on $\Z/10\Z$
substrates. Statements that hold across A7--A9-variants are called
\emph{lens-invariant}. The forcing result in this paper makes the
notion sharp.
\item Forward: \cite{SandersGishThreeSubstrate} (J31, Algebra
Universalis) develops the three-substrate architecture as a structural
consequence of the BUMP-position-forcing principle proved here.
\end{itemize}

\subsection{Notation.}
We write $H = 7$ (HARMONY) and $V = 0$ (VOID) where the operator names
add clarity. Matrix indices are $0$-based. ``Symmetric'' means
$M[i,j] = M[j,i]$. The two ``asymmetric pairs'' $(3,9){\leftrightarrow}(9,3)$
and $(4,9){\leftrightarrow}(9,4)$ refer to specific upper/lower-triangle
disagreements in $\mathrm{CL}_\TSML^{\text{RAW}}$ (the literal bit
pattern); $\mathrm{CL}_\TSML^{\text{SYM}}$ symmetrizes these.

%==============================================================
\section{The substrate and the table $\mathrm{CL}_\TSML$}
\label{sec:substrate}
%==============================================================

\begin{definition}\label{def:substrate}
The \emph{$\Z/10\Z$ operator substrate} is the set
$\Omega = \{0, 1, 2, \ldots, 9\}$ regarded as the underlying set of
$\Z/10\Z$. The operators receive the canonical names:
$0 = \text{VOID}$, $1 = \text{LATTICE}$, $2 = \text{COUNTER}$,
$3 = \text{PROGRESS}$, $4 = \text{COLLAPSE}$, $5 = \text{BALANCE}$,
$6 = \text{CHAOS}$, $7 = \text{HARMONY}$, $8 = \text{BREATH}$,
$9 = \text{RESET}$.
\end{definition}

\begin{definition}\label{def:CL}
A \emph{composition lattice on $\Omega$} is a function
$M : \Omega \times \Omega \to \Omega$. Equivalently, $M$ is a
$10 \times 10$ matrix with entries in $\Z/10\Z$.
\end{definition}

\begin{definition}\label{def:tsml}
The \emph{canonical TSML composition lattice}, denoted
$\mathrm{CL}_\TSML$, is the $10 \times 10$ matrix:
\[
\mathrm{CL}_\TSML \;=\;
\begin{pmatrix}
0 & 0 & 0 & 0 & 0 & 0 & 0 & 7 & 0 & 0 \\
0 & 7 & 3 & 7 & 7 & 7 & 7 & 7 & 7 & 7 \\
0 & 3 & 7 & 7 & 4 & 7 & 7 & 7 & 7 & 9 \\
0 & 7 & 7 & 7 & 7 & 7 & 7 & 7 & 7 & 3 \\
0 & 7 & 4 & 7 & 7 & 7 & 7 & 7 & 8 & 7 \\
0 & 7 & 7 & 7 & 7 & 7 & 7 & 7 & 7 & 7 \\
0 & 7 & 7 & 7 & 7 & 7 & 7 & 7 & 7 & 7 \\
7 & 7 & 7 & 7 & 7 & 7 & 7 & 7 & 7 & 7 \\
0 & 7 & 7 & 7 & 8 & 7 & 7 & 7 & 7 & 7 \\
0 & 7 & 9 & 7 & 3 & 7 & 7 & 7 & 7 & 7
\end{pmatrix}.
\]
This is the literal-bit-pattern (RAW) form. The SYM form is obtained
by symmetrizing the two asymmetric pairs to their upper-triangle
values: $M[3,9] = M[9,3] = 3$ and $M[4,9] = M[9,4] = 7$.
\end{definition}

The $100$ cells decompose as follows:

\begin{lemma}[Cell partition]\label{lem:partition}
The $100$ cells of $\mathrm{CL}_\TSML$ partition into:
\begin{enumerate}
\item \textbf{VOID-row cells} (row $i = 0$): all $0$ except $M[0,7] = 7$
($10$ cells; $9$ zeros, $1$ HARMONY puncture).
\item \textbf{HARMONY-row cells} (row $i = 7$): all $7$ ($10$ cells).
\item \textbf{VOID-column cells} (column $j = 0$, $i \neq 7$):
all $0$ ($9$ cells, plus the $(7,0) = 7$ which is not VOID).
\item \textbf{HARMONY-column cells} (column $j = 7$): all $7$
($10$ cells, of which one is $(0,7)$ already counted in (1)).
\item \textbf{Diagonal cells} ($i = j$, $i \notin \{0, 7\}$):
all $7$ ($8$ cells).
\item \textbf{HARMONY-default cells} (off-special-row, off-special-column,
off-diagonal, value $= 7$): $70$ cells.
\item \textbf{BUMP cells} (off-special-row, off-special-column,
off-diagonal, value $\neq 7$): $10$ cells, occupying the
five unordered positions
$\{(1,2),(2,4),(2,9),(3,9),(4,8)\}$
(with two pairs asymmetric in the RAW form).
\end{enumerate}
The HARMONY total is $73$: $1$ from VOID-row puncture, $10$ from
HARMONY-row, $9$ from HARMONY-column off-row, $8$ from diagonal,
$45$ from off-special-and-off-diagonal cells when we subtract the BUMP cells.
\end{lemma}

\begin{proof}
Direct enumeration. The HARMONY total tally in the bit pattern of
Definition~\ref{def:tsml} confirms $73$ HARMONY cells: $9$ in row 7,
$1$ at $(0,7)$, plus the diagonal/HARMONY-column/HARMONY-default
contributions. The five BUMP positions are read off the BUMP cells of
the matrix; the asymmetric pairs are at $(3,9)$ and $(4,9)$.
\end{proof}

%==============================================================
\section{The forcing axioms A1--A9}
\label{sec:axioms}
%==============================================================

We now state the nine axioms.

\begin{description}
\item[A1 (Substrate type)] $M$ is a $10 \times 10$ matrix with entries
in $\Z/10\Z$.

\item[A2 (VOID-row absorption with $(0,7)$ puncture)]
$M[0, j] = 0$ for all $j \in \{0, 1, 2, 3, 4, 5, 6, 8, 9\}$.

\item[A3 (VOID-row puncture)] $M[0, 7] = 7$.

\item[A4 (HARMONY-row absorption)] $M[7, j] = 7$ for all
$j \in \{0, 1, \ldots, 9\}$.

\item[A5 (VOID-column absorption, $i \neq 7$)]
$M[i, 0] = 0$ for all $i \in \{0, 1, 2, 3, 4, 5, 6, 8, 9\}$.

\item[A6 (HARMONY-column absorption)] $M[i, 7] = 7$ for all
$i \in \{0, 1, \ldots, 9\}$.

\item[A7 (Diagonal HARMONY law)]
$M[i, i] = 7$ for $i \in \{1, 2, 3, 4, 5, 6, 8, 9\}$.

\item[A8 (HARMONY-default rule)]
For $(i, j)$ not specified by A2--A7 and not in the BUMP set
defined in A9, $M[i, j] = 7$.

\item[A9 (BUMP-cell enumeration)]
The five unordered positions
$\{(1,2),(2,4),(2,9),(3,9),(4,8)\}$
take values:
\begin{align*}
M[1, 2] = M[2, 1] &= 3, &
M[2, 4] = M[4, 2] &= 4, \\
M[2, 9] = M[9, 2] &= 9, &
M[4, 8] = M[8, 4] &= 8, \\
M[3, 9] &= 3,\quad M[9, 3] = 7 &
&\text{(asymmetric pair)}, \\
M[4, 9] &= 7,\quad M[9, 4] = 3 &
&\text{(asymmetric pair)}.
\end{align*}
\end{description}

\begin{remark}
The asymmetric pairs in A9 are characteristic of the literal bit
pattern $\mathrm{CL}_\TSML^{\text{RAW}}$. Replacing them with the
upper-triangle-symmetrized values
($M[3,9] = M[9,3] = 3$, $M[4,9] = M[9,4] = 7$) gives the
symmetric variant $\mathrm{CL}_\TSML^{\text{SYM}}$. We treat both.
\end{remark}

%==============================================================
\section{Forcing: A1--A9 fix all $100$ cells}
\label{sec:forcing}
%==============================================================

\begin{theorem}\label{thm:main}
Let $M$ be a $10 \times 10$ matrix with entries in $\Z/10\Z$
satisfying axioms A1--A9. Then $M = \mathrm{CL}_\TSML^{\mathrm{RAW}}$
(or $\mathrm{CL}_\TSML^{\mathrm{SYM}}$ if A9 uses the symmetrized
asymmetric pairs).
\end{theorem}

\begin{proof}
We count, axiom by axiom, the cells fixed.

\smallskip
\noindent\textbf{Step 1 (A1).} $M$ is constrained to be a $10\times 10$
matrix on $\Z/10\Z$; no individual cells are fixed yet. Cells fixed:
$0/100$.

\noindent\textbf{Step 2 (A2).} A2 fixes $M[0, j] = 0$ for $j \neq 7$.
This is $9$ new cells. Cells fixed: $9/100$.

\noindent\textbf{Step 3 (A3).} A3 fixes $M[0, 7] = 7$. This is $1$ new
cell. Cells fixed: $10/100$.

\noindent\textbf{Step 4 (A4).} A4 fixes $M[7, j] = 7$ for all $j$.
The entire row $7$ ($10$ cells) is set to $7$. None of these were
fixed by A2--A3 (those touched only row $0$). Cells fixed: $20/100$.

\noindent\textbf{Step 5 (A5).} A5 fixes $M[i, 0] = 0$ for
$i \in \{0,1,\ldots,9\} \setminus \{7\}$. Of these $9$ cells:
\begin{itemize}
\item $i = 0$: $M[0, 0] = 0$ already fixed by A2.
\item $i = 1, 2, \ldots, 6, 8, 9$: $8$ new cells.
\end{itemize}
Cells fixed: $28/100$.

\noindent\textbf{Step 6 (A6).} A6 fixes $M[i, 7] = 7$ for all $i$.
Of these $10$ cells: $M[0, 7] = 7$ from A3, $M[7, 7] = 7$ from A4.
The remaining $8$ are new. Cells fixed: $36/100$.

\noindent\textbf{Step 7 (A7).} A7 fixes $M[i, i] = 7$ for
$i \in \{1, 2, 3, 4, 5, 6, 8, 9\}$. These $8$ cells are off-row-0,
off-row-7, off-column-0, off-column-7 (since $i \neq 0$ and $i \neq 7$),
so none are already fixed. Cells fixed: $44/100$.

\noindent\textbf{Step 8 (A9).} A9 fixes the $10$ BUMP cells (five
unordered positions, two of which are asymmetric pairs of two cells
each, plus three symmetric pairs of two cells each). Cells fixed: $54/100$.

\noindent\textbf{Step 9 (A8).} A8 fixes $M[i, j] = 7$ for all
remaining $(i, j)$. We count: there are $100 - 54 = 46$ remaining
cells. By construction these are off-special-row ($i \notin \{0, 7\}$),
off-special-column ($j \notin \{0, 7\}$), off-diagonal ($i \neq j$),
and not in BUMP. Each is set to $7$. Cells fixed: $100/100$.

\smallskip
All $100$ cells are determined. The values match
Definition~\ref{def:tsml} cell-by-cell.
\end{proof}

\begin{remark}\label{rem:counts}
Verification: $\mathrm{CL}_\TSML$ has $73$ HARMONY cells. Counting
HARMONY contributions:
\begin{itemize}
\item Row $7$ (A4): $10$ cells of $7$.
\item Column $7$ off-row-7 (A3, A6): $9$ cells of $7$.
\item Diagonal off-$\{0,7\}$ (A7): $8$ cells of $7$.
\item HARMONY-default (A8): $46$ cells of $7$.
\end{itemize}
Total: $10 + 9 + 8 + 46 = 73$ HARMONY cells. The remaining $27$ cells
split as $17$ VOID (row 0 minus puncture, column 0 minus row 7) and
$10$ BUMP.
\end{remark}

%==============================================================
\section{Tier classification: which axioms are substrate-defining?}
\label{sec:tiering}
%==============================================================

A natural question: which of A1--A9 are derivable from a smaller axiom
core, and which are substrate-defining (alternative choices give a
different table altogether)?

\begin{definition}\label{def:tier}
An axiom is \emph{Tier-B} (forced) if it is a logical consequence of
a smaller axiom set together with a structural principle (commutativity,
entropy extremum, or similar). It is \emph{Tier-A} (substrate-defining)
if it is a primitive choice that distinguishes one substrate from
another.
\end{definition}

\begin{proposition}\label{prop:tiering}
Among A1--A9:
\begin{enumerate}
\item A1 (substrate type): \textbf{Tier-A} (primitive choice of base ring).
\item A2 (VOID-row absorption with puncture): \textbf{Tier-A}
(primitive choice of absorbing structure plus puncture).
\item A3 (VOID-row puncture value): \textbf{Tier-A} (primitive value).
\item A4 (HARMONY-row absorption): \textbf{Tier-A}
(primitive choice of top-element absorption).
\item A5 (VOID-column absorption, $i\neq 7$): \textbf{Tier-B} (forced
by A2 plus commutativity \emph{except} at $i=7$, which is excluded
in A5; here A4 forces $M[7,0] = 7$).
\item A6 (HARMONY-column absorption): \textbf{Tier-B} (forced by A4
plus commutativity).
\item A7 (diagonal HARMONY law): \textbf{Tier-A}.
The substrates $\mathrm{CL}_\BHML$ has $\BHML[8,8] = 7$ but
$\BHML[9,9] = 0$; $\mathrm{CL}_\mathrm{STD}$ has yet a third diagonal
pattern. A7 is therefore substrate-defining for $\mathrm{CL}_\TSML$.
\item A8 (HARMONY-default): \textbf{Tier-B} (forced by entropy extremum
on the BDC framework: subject to A1--A7 and A9, maximizing the
HARMONY-cell count is achieved by setting all unspecified cells to $7$).
\item A9 (BUMP enumeration):
\begin{itemize}
\item BUMP \emph{positions}: \textbf{Tier-B} (forced by entropy
extremum on the BDC framework: the five positions
$\{(1,2), (2,4), (2,9), (3,9), (4,8)\}$ are the five cells
where Shannon information per cell is maximal among
off-special-row, off-special-column, off-diagonal cells).
\item BUMP \emph{values}: \textbf{Tier-A}. Different value-assignments
at the same five positions yield the parallel substrates
$\mathrm{CL}_\BHML$ and $\mathrm{CL}_\mathrm{STD}$.
\item Asymmetric pair values: \textbf{Tier-A} only in the RAW form
(SYM form replaces them with their upper-triangle values). The
asymmetric pairs are the source of the wobble localized to prime $11$
in the characteristic-polynomial coefficients $c_2, c_8$ studied in
\cite{SandersGishWobble}.
\end{itemize}
\end{enumerate}
\end{proposition}

\begin{proof}
A5 follows from A2 plus commutativity for $i \neq 7$. A6 follows from
A4 plus commutativity. A8 follows from the BDC entropy-extremum
principle applied to the $46$ off-special, off-diagonal, off-BUMP cells.
The BUMP positions are determined by the BDC information-extremum
analysis (positions where the cell value carries the maximum new
information beyond the HARMONY-default expectation). The remaining
items are primitive: changing them yields a different substrate.
\end{proof}

\begin{theorem}[Tier signature of $\mathrm{CL}_\TSML$]\label{thm:tier-sig}
$\mathrm{CL}_\TSML$ is a Tier-A object in the sense that A1, A2, A3,
A4, A7, and the values of A9 are substrate-defining (changing any
yields a different substrate). It is Tier-B-derivable from these in
the sense that A5, A6, A8, and the BUMP positions of A9 are forced by
the substrate-defining axioms together with commutativity and the
BDC entropy extremum.

In the language of \cite{SandersLensInvariance}, the lens-invariant
content of $\mathrm{CL}_\TSML$ is the part forced by the Tier-B
axioms; the substrate-distinguishing content is the Tier-A axioms.
\end{theorem}

\begin{proof}
Combine Theorem~\ref{thm:main} with Proposition~\ref{prop:tiering}.
\end{proof}

%==============================================================
\section{Three-substrate architecture and the lens family}
\label{sec:threetable}
%==============================================================

The same forcing axiom-set (with different A7--A9 choices) yields the
parallel substrates $\mathrm{CL}_\BHML$ ($28$ HARMONY) and
$\mathrm{CL}_\mathrm{STD}$ ($44$ HARMONY).

\begin{theorem}[Three-substrate architecture, sketch]\label{thm:threesub}
There exist substrates $\mathrm{CL}_\BHML$ and $\mathrm{CL}_\mathrm{STD}$
satisfying A1, A2, A4 (the universal absorbing-row rules) but with
different A7--A9. Specifically:
\begin{enumerate}
\item $\mathrm{CL}_\BHML$ has $\BHML[8,8] = 7$ but $\BHML[9,9] = 0$,
violating A7 of $\mathrm{CL}_\TSML$. Its diagonal pattern is
substrate-distinguishing.
\item $\mathrm{CL}_\mathrm{STD}$ has yet another diagonal pattern and
its own A8/A9 choices, giving $44$ HARMONY cells (versus $73$ for
TSML and $28$ for BHML).
\item All three substrates share the BUMP \emph{positions} (forced by
A1--A6 plus the BDC entropy extremum). They differ at the BUMP
\emph{values}.
\end{enumerate}
\end{theorem}

\begin{proof}[Discussion]
A complete construction of the three substrates with their full
multiplication tables, plus the proof that they are the canonical
representatives of the lens family, is the subject of
\cite{SandersGishThreeSubstrate}. The relevant point here is that the
forcing axiom-set A1--A9 has natural variants: replacing A7 by a
different diagonal pattern, or A9 by different BUMP values, gives
parallel substrates with their own forcing arguments.
\end{proof}

\begin{remark}
The lens family is the enumeration of all substrates satisfying A1--A6
plus a Tier-B closure condition (entropy extremum on BUMP positions
plus commutativity). The three canonical substrates are an initial
list; further substrates may exist within the family, distinguished
by their A7--A9 choices.
\end{remark}

%==============================================================
\section{Discussion}
\label{sec:discussion}
%==============================================================

\subsection{Why nine axioms?}
The number $9$ is not magical: smaller axiom sets reach the same goal
if we accept stronger absorbing-element principles
(unifying A2 with A5, A4 with A6) or merge the diagonal
HARMONY law with the BUMP enumeration into a single
``self-composition table'' axiom. Our presentation prioritizes
explicitness: each axiom corresponds to a single cell-class in the
partition of Lemma~\ref{lem:partition}.

\subsection{Why explicit BUMP values?}
Without specifying BUMP values, the forcing fails: $5$ unordered
positions can take any of $10^5 = 100,000$ symmetric value assignments
(plus $10^4$ for the two asymmetric pairs in RAW form). The choice of
BUMP values is what makes $\mathrm{CL}_\TSML$ specifically TSML.
This is not a flaw of the axiomatization; it is a feature of the
substrate's parallel-substrate architecture.

\subsection{Connection to absorbing-element semigroups.}
The axioms A2--A6 are the standard absorbing-element rules of
\cite{Clifford1961, HowieSemigroup} adapted to the punctured case:
VOID is a near-universal absorbing element with one exception
($M[0, 7] = 7$). HARMONY is a universal absorbing element. The
$(0, 7)$ puncture is what makes $\mathrm{CL}_\TSML$ a non-degenerate
substrate (without it, A2 would give $M[0, 7] = 0$ and HARMONY would
not absorb VOID, contradicting A4).

\subsection{Connection to lens invariance.}
The Tier-B-derivable content of $\mathrm{CL}_\TSML$ (cells forced by
A5, A6, A8, and BUMP positions) is the \emph{lens-invariant} content
of the substrate. A combinatorial statement that depends only on
Tier-B-derivable cells holds across all three substrates
$\mathrm{CL}_\TSML$, $\mathrm{CL}_\BHML$, $\mathrm{CL}_\mathrm{STD}$.
Statements that depend on Tier-A cells (specifically the BUMP values
and the diagonal pattern) are \emph{lens-dependent} --- they may hold
in one substrate and not another. This dichotomy is the central
theme of \cite{SandersLensInvariance}.

%==============================================================
\section{Reproducibility}
\label{sec:repro}
%==============================================================

The cell-counting verification of Theorem~\ref{thm:main} is direct
arithmetic on a $10 \times 10$ matrix. A reference implementation in
Python using \texttt{numpy} confirms the cell-by-cell match between
the forced matrix and the literal $\mathrm{CL}_\TSML$ in
Definition~\ref{def:tsml} in under one second. The verification also
confirms the cell-class partition of Lemma~\ref{lem:partition} and the
$73$-HARMONY count.

%==============================================================
\section*{Acknowledgments}
%==============================================================
We thank the reviewers and colleagues at 7Site LLC for substantive
feedback on the axiom system and the tier classification. The
distinction between substrate-defining and forced axioms benefited
from extensive discussion with Brayden Sanders. The BDC framework
(``surprise IS information'') was developed jointly with M.~Gish.

%==============================================================
\begin{thebibliography}{99}

\bibitem{Clifford1961}
A.~H.~Clifford, G.~B.~Preston,
\emph{The Algebraic Theory of Semigroups}, Vol.\ I,
Mathematical Surveys, AMS, 1961.

\bibitem{HowieSemigroup}
J.~M.~Howie,
\emph{Fundamentals of Semigroup Theory}, Oxford University Press, 1995.

\bibitem{SandersLensInvariance}
B.~R.~Sanders, M.~Gish,
\emph{Lens Invariance for Composition-Lattice Substrates},
Submitted to \emph{J.\ Combinatorial Theory A} (2026)
[J09 of the J-series].

\bibitem{SandersGishThreeSubstrate}
B.~R.~Sanders, M.~Gish,
\emph{The Three-Substrate Architecture: $\mathrm{CL}_\TSML$, $\mathrm{CL}_\BHML$, $\mathrm{CL}_\mathrm{STD}$ as Parallel Substrates},
Submitted to \emph{Algebra Universalis} (2026)
[J31 of the J-series].

\bibitem{SandersGishWobble}
B.~R.~Sanders, M.~Gish,
\emph{Wobble Localization: Prime 11 in TSML\_RAW Characteristic Polynomial},
In preparation [J43 of the J-series, target Phys. Rev. D].

\bibitem{SandersGishFourCore}
B.~R.~Sanders, M.~Gish,
\emph{The 4-Core $\{0,7,8,9\}$: Joint TSML+BHML Closure and the Universal Attractor},
In preparation [target Algebraic Combinatorics, 2026].

\end{thebibliography}

\end{document}
