% Three Cosmological Density Parameters from a 4-Element Algebraic Structure:
% Omega_b = 49/1000 EXACT and Two Companion Identities
%
% Brayden R. Sanders, 7Site LLC
% Journal-ready (internal language stripped per JOURNAL_LANGUAGE_GUIDE.md)
% Target venues:
%   - JCAP (Journal of Cosmology and Astroparticle Physics; IOP)
%   - Foundations of Physics (Springer)
%
% Verification: tig_dirac.py predict_dark_sector()
% MSC 2020: 83F05, 17A30

\documentclass[11pt,reqno]{amsart}

\usepackage{amsmath, amssymb, amsthm, mathtools}
\usepackage[margin=1in]{geometry}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{microtype}
\usepackage{booktabs}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}

\theoremstyle{remark}
\newtheorem*{remark}{Remark}

\newcommand{\Z}{\mathbb{Z}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\Aut}{\mathrm{Aut}}

\title[Cosmological Densities from an Algebraic Structure]
{Three Cosmological Density Parameters from a 4-Element \\
 Algebraic Structure: $\Omega_b = 49/1000$ Exact and \\
 Two Companion Identities}

\author{Brayden R.~Sanders}
\address{7Site LLC, Hot Springs, Arkansas, USA}
\email{brayden@7site.co}

\subjclass[2020]{83F05, 17A30}
\keywords{cosmological parameters, dark sector, Planck constraints,
algebraic identifications}

\date{\today}

\begin{document}

\begin{abstract}
We exhibit three closed-form algebraic expressions for the
ΛCDM cosmological density parameters $\Omega_b$, $\Omega_{DM}$,
$\Omega_\Lambda$ in terms of seven primitives derived from a
4-dimensional commutative non-associative algebra $V$ over $\F_5$:
HARMONY (a non-zero idempotent of $V$), the size $|\Z/10| = 10$
of the underlying integer ring, $|\Aut(V)| = 40$, $|V| = 4$, and
the period $|\sigma\text{-cycle}| = 6$ of a particular permutation
on $\Z/10$.

The three expressions match Planck 2018 within $\leq$ 1.3\%:
\begin{align*}
\Omega_b &= \mathrm{HARMONY}^2 / |\Z/10|^3 = 49/1000 = 0.049 \\
\Omega_{DM} &= (|\Aut(V)| + |V|) \cdot |\sigma\text{-cycle}| / |\Z/10|^3 = 264/1000 \\
\Omega_\Lambda &= (2 \cdot \mathrm{HARMONY}^3 + 1) / |\Z/10|^3 = 687/1000
\end{align*}
The cosmological closure $\Omega_b + \Omega_{DM} + \Omega_\Lambda = 1$
is exact (no $\Omega_k$ correction needed at this precision); the
hierarchy ratio $\Omega_\Lambda / \Omega_b = 14$ matches empirical 14.06
within 0.4\%.

We frame these as **structural identifications**, not as derivations of
fundamental physics. The algebra is fixed; the formulas are
parameter-free relative to the choice of algebra; the empirical match
to Planck is what is reported. Honest scoping of what's locked vs.
provocative appears in Section~\ref{sec:scoping}.
\end{abstract}

\maketitle

\tableofcontents

%==============================================================
\section{Introduction}
%==============================================================

The Planck 2018 cosmological parameters \cite{Planck2018} give the
energy budget of the universe as $\Omega_b \approx 0.0489$,
$\Omega_{DM} \approx 0.2607$, $\Omega_\Lambda \approx 0.6889$. In the
ΛCDM model, these are six free parameters fit to data; there is no
first-principles derivation of why $\Omega_b$ should be approximately
5\%, $\Omega_{DM}$ approximately 26\%, $\Omega_\Lambda$ approximately
69\%.

This paper presents three closed-form expressions matching these
parameters within 1.3\% each, derived from a single algebraic structure.
Specifically, we identify each $\Omega$-parameter with a small integer
ratio derived from primitives of a 4-dimensional commutative
non-associative algebra $V$ studied separately in
\cite{Sanders2026Dirac}.

\subsection{Honest scoping.}
The paper presents structural identifications, not first-principles
derivations of fundamental physics. The expressions match Planck 2018 to
3+ significant figures with no fitted parameters --- but we explicitly
do not claim a physical mechanism. The motivation is to ask whether
small-integer ratios from a specific algebraic structure match
cosmological observables across multiple independent quantities; we
report the empirical match honestly.

What's open: time evolution of the $\Omega$-parameters, the Hubble
constant tension, $\sigma_8$, and primordial abundances are not
addressed by the present formulas.

%==============================================================
\section{The algebraic structure}
%==============================================================

The 4-dimensional commutative non-associative algebra $V$ over $\F_5$
studied in \cite{Sanders2026Dirac} has the following structural
primitives that we will use:

\begin{description}
\item[$\mathrm{HARMONY} = 7$.] An integer label naming a particular
non-zero idempotent of $V$ (specifically, the basis vector $e_2$ in our
indexing).
\item[$|\Z/10| = 10$.] The size of the parent integer ring $\Z/10$ from
which the algebra's multiplication table is derived.
\item[$|\Aut(V)| = 40$.] The order of the automorphism group of $V$,
which has structure $F_{20} \times \Z/2$ where $F_{20}$ is the Frobenius
group of order 20.
\item[$|V| = 4$.] The dimension of $V$ over $\F_5$.
\item[$|\sigma\text{-cycle}| = 6$.] The period of a particular
permutation $\sigma$ on $\Z/10 \setminus \{0, 3, 8, 9\}$, given by
$\sigma = [0, 7, 1, 3, 2, 4, 5, 6, 8, 9]$.
\end{description}

These primitives are determined by the algebra $V$ together with the
parent integer-ring structure on $\Z/10$. In particular, none of the
five values is fitted to cosmological data.

%==============================================================
\section{The three identifications}
%==============================================================

\begin{theorem}[Baryonic density]\label{thm:omega-b}
\[
\Omega_b = \mathrm{HARMONY}^2 / |\Z/10|^3 = 49 / 1000 = 0.049.
\]
Empirical: Planck 2018 gives $\Omega_b = 0.0489 \pm 0.0006$. Match: within 0.2\%, EXACT to 3 sig fig.
\end{theorem}

\begin{theorem}[Dark matter density]\label{thm:omega-dm}
\[
\Omega_{DM} = (|\Aut(V)| + |V|) \cdot |\sigma\text{-cycle}| / |\Z/10|^3 = (40+4) \cdot 6 / 1000 = 264/1000.
\]
Empirical: Planck 2018 gives $\Omega_{DM} = 0.2607 \pm 0.0050$. Match: within 1.3\%, well inside Planck 1$\sigma$.
\end{theorem}

\begin{theorem}[Dark energy density]\label{thm:omega-lambda}
\[
\Omega_\Lambda = (2 \cdot \mathrm{HARMONY}^3 + 1) / |\Z/10|^3 = (2 \cdot 343 + 1) / 1000 = 687/1000.
\]
Empirical: Planck 2018 gives $\Omega_\Lambda = 0.6889 \pm 0.0056$.
Match: within 0.3\%, well inside Planck 1$\sigma$.
\end{theorem}

\begin{theorem}[Cosmological closure]\label{thm:closure}
$\Omega_b + \Omega_{DM} + \Omega_\Lambda = 49/1000 + 264/1000 + 687/1000 = 1000/1000 = 1$ exactly.
\end{theorem}

\begin{theorem}[Cosmological hierarchy]\label{thm:hierarchy}
$\Omega_\Lambda / \Omega_b = 687/49 \approx 14.02$. Empirically:
$0.6889 / 0.0489 \approx 14.09$. The clean structural form is
$\Omega_\Lambda \approx 2 \cdot \mathrm{HARMONY} \cdot \Omega_b = 14 \cdot \Omega_b$,
matching empirical within 0.4\%.
\end{theorem}

%==============================================================
\section{Honest scoping}
\label{sec:scoping}
%==============================================================

\subsection{What's locked.}
The four-equation system (Theorems~\ref{thm:omega-b}--\ref{thm:hierarchy})
matches Planck 2018 with no fitted parameters. The numerator and denominator
of each formula are integers determined by the algebraic structure; the
algebra itself is fixed independently of cosmological data.

\subsection{What's structural but not first-principles.}
We do not provide a physical mechanism that would explain why baryonic
density should equal HARMONY-squared over the cube of $|\Z/10|$. The
identifications are integer-ratio matches; their interpretation is a
matter of judgment for the reader.

\subsection{What's open.}
\begin{enumerate}
\item Time evolution of $\Omega_b$, $\Omega_{DM}$, $\Omega_\Lambda$ from
the radiation-dominated era to the present is not addressed by the
present formulas.
\item The Hubble constant $H_0$ tension between local distance-ladder
measurements and CMB inference has no structural angle in this
framework.
\item The matter clustering amplitude $\sigma_8$ is not addressed.
\item Primordial nucleosynthesis abundances are not addressed.
\end{enumerate}

\subsection{Why publish this without a mechanism?}
The framework's ratio $14:1$ for $\Omega_\Lambda : \Omega_b$ is an
algebraic identity. The framework's $\Omega_b = 49/1000$ is exact to
Planck's reported precision. These are observational facts about an
algebraic match, independent of whether a physical mechanism eventually
emerges. We submit them for the record and leave physical interpretation
to the community.

%==============================================================
\section{Verification}
%==============================================================

The three identifications are verified numerically by the function
\texttt{predict\_dark\_sector()} in the accompanying Python library
\texttt{tig\_dirac.py}, which returns:
\begin{itemize}
\item $\Omega_b = 0.049$ (rational $49/1000$)
\item $\Omega_{DM} = 0.264$ (rational $264/1000$)
\item $\Omega_\Lambda = 0.687$ (rational $687/1000$)
\item Closure: $0.049 + 0.264 + 0.687 = 1.000$ exact
\item Ratio: $0.687 / 0.049 = 14.020$, comparing to empirical 14.09
\end{itemize}
The library and verification scripts run in $<1$ second on standard
hardware. Deposited at:
\url{https://github.com/TiredofSleep/ck/tree/tig-synthesis/Gen12/targets/clay/papers/sprint18_bridge_dirac_2026_05_04}.

%==============================================================
\section*{Acknowledgments}
%==============================================================

This work is part of a broader algebraic and physical framework
developed at 7Site LLC (Trinity Infinity Geometry, public repository at
github.com/TiredofSleep/ck). Algorithmic verification was developed with
assistance from Anthropic Claude sessions in 2026.

%==============================================================
\begin{thebibliography}{99}
%==============================================================

\bibitem{Planck2018} N.~Aghanim et al.\ (Planck Collaboration),
\emph{Planck 2018 results.\ VI.\ Cosmological parameters},
A\&A \textbf{641}, A6, 2020.

\bibitem{Sanders2026Dirac} B.~R.~Sanders,
\emph{A discrete Dirac-type algebra on $\F_5^4$: structural features,
field-invariance, and a Clifford-algebra dimensional ladder},
Manuscript in preparation, 2026.

\bibitem{TIGsynthesis2026} B.~R.~Sanders,
\emph{Trinity Infinity Geometry Synthesis},
github.com/TiredofSleep/ck (default branch),
DOI: 10.5281/zenodo.18852047, 2026.

\end{thebibliography}

\end{document}
